# Preliminary Content {-}

## Acknowledgements {-}

I want to thank a few people.

## Preface {-}

This is an example of a thesis setup to use the reed thesis document class
(for LaTeX) and the R bookdown package, in general.

## Dedication {-}

You can have a dedication here if you wish. 

`r if(!knitr:::is_latex_output()) '## Abstract {-}'`





<!--
Had trouble getting this to supress labeling with Chapter 1, etc.
Tried to use '{-}' like in the reference section.
-->



<!--
Numbering
All pages in the document should be numbered, with the exception of the title page. Use the Roman numeral i (one) for the first page after the title page. 
Use the Arabic numerical number 1 (one) for the first page of the body. 
Upper case letters should be used to number appendixes. 
Page numbers must be in the same font and size as the body of the text and placed on all pages, including the Abstract and the Table of Contents. 
Page numbers must be in the upper right-hand corner, 3⁄4 inch down from the top, and 3⁄4 inch from the right-hand side.

Margins
• 1” margin on the top, bottom, and right (includes page number)
• 1.5” margin on the left, including the title page

The same margins apply to both landscape and portrait page orientation. 
Margins are to be blank—without punched holes, writing, or other marks. The margins set in many word processing packages may not measure the same on the printed page. Measure the margins on printed test pages and adjust the software settings accordingly to meet the above requirements.



Pages which have printed text only must be oriented in the same manner as the pages of this instruction manual, often called “portrait.” When necessary, graphs, photographs, tables, and figures may be oriented facing to the right. The manuscript is to be double- spaced with paragraph indentation and tables typed in accordance with format of journals in the candidate’s particular discipline and printed on one side of the paper only.
Style comments:
All pages must have the following format:



• Print on one side of the page only (no duplex printing)
• Page numbers in the upper right hand corner, 3⁄4” from the top and right side. No page number on the title page.

Body text: 12 pt minimum font size, double-spaced, 

Recommended font is Times New Roman for the body text 
Fonts used should be clean and easily readable. The same font should be used for the principle text throughout the work, although other fonts and styles may be employed for smaller segments.

Figure/Table captions: 
Recommended font is Times New Roman, Calibri or Helvetica for the headings.

12 pt minimum font (10 pt minimum for captions) 10 pt minimum font size, single or double-spaced, must be distinctive in size or style from the body text.

Only the first word of each sentence of captions should be capitalized. Add 1 or 2 empty lines after captions to enable readers to readily differentiate captions from narrative text.

Figures, maps, graphs and illustrations may be left-justified or center-justified. Captions for all figures, maps, graphs and illustrations must be within margin and font limitations.
All figure captions must be placed at the bottom or on the right-hand side of an illustration and left-justified. 
Captions should be in a different font to distinguish may be single- or double-spaced and should be distinctive from the body text in size (10-pt minimum) or style from the body of text. 


Use of Acronyms:
At the first usage of an acronym spell out entire name/phrase for which an acronym is to be used and follow with the acronym in parentheses. Avoid off-color or otherwise inappropriate acronyms (which removes ~5% of the fun of writing)



The bibliographic style used in the thesis or dissertation is determined by the candidate’s department or professional society and by consultation with the candidate’s major professor. 

ASCE Style Guide:
http://ascelibrary.org/doi/pdf/10.1061/9780784478998.ch11 

The Office of Graduate Education encourages preparation of theses/dissertations in a format suitable for submission for publication in leading journals in the candidate’s field.

The following section provides examples for citing a wide variety of sources in the American Psychological Association (APA) style, a commonly used style similar to the bibliographic styles used in many disciplines. In some fields, styles may differ considerably. The candidate should follow a format required by the publication to which he/she plans to submit the work. However, if there is no specific format in a given field, and the candidate’s department has no specific guidelines, use of this APA style is encouraged; however, every discipline has its own protocols and students may, with the consent of their committee, use the style guidelines of their discipline.

The author may use either footnotes or references to a bibliography at the end of the thesis or dissertation. As stated above, bibliographic style (footnotes, references, citations, etc.) and type (mathematical notation, symbols, etc.) should be those consistent with the guidelines established by the candidate’s department or professional society. 
The style of the principal abstracting journal of the candidate’s discipline is otherwise a good guide. Whenever possible, periodical titles should be abbreviated according to the standards outlined by the American National Standards Institute.

Some specific journals and bibliographies in various disciplines are:
American National Standards Institute, Inc., American National Standard for the Abbreviation of Titles of Periodicals. 1970. (Z39.5-1969—which is a revision of Z39.5-1963).


Civil Engineering: ASCE Publications Abstracts; Engineering Index. 
-->

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

# Title & Signature Page 
<!--(required; considered in numbering but NOT numbered)
Specific format required; see pages 13 & 14 for example
Needs major work to get formatting correct!
Currently below the TOC, needs to be above the TOC
Also needs centered, etc.

The title page must match the template provided in this instruction manual for the year you are completing your degree, and must contain the title, degree program, author, whether the document is a master’s thesis or a doctoral dissertation, its purpose, name of the college, date defended, and signature approvals of the Graduate Advisory Committee, the head of the major department or program coordinator of a multi- disciplinary program, and the Dean of Graduate Education. Names should be printed below signatures. Examples of M.S. and PhD. title pages follow.

For the signature lines of SD Mines faculty committee members, there are some cases where the faculty member’s departmental affiliation does not fully express their close affiliation with an interdisciplinary graduate program. In such cases it is acceptable to list both the faculty member’s home department and their interdisciplinary affiliation. SD Mines has four interdisciplinary graduate programs: Materials Engineering and Science, Biomedical Engineering, Atmospheric and Environmental Sciences and Nanoscience and Nanoengineering.

For the signature lines of external committee members, their institutional affiliation (if any) should be listed.
In all cases the committee member’s highest degree should be listed.

Note that 12 point font is the required standard for theses and dissertations. If you are unable to fit your committee signatures onto one page, please treat the position titles as captions and reduce the font to 10 points.
-->

The Metaphysics and Mythology of Coffee  
by  
Joseph Highly Caffeinated  
A thesis submitted to the Graduate Division  
in partial fulfillment of the requirements for the degree of  
Master of Science in Chemical Engineering  
South Dakota School of Mines and Technology Rapid City, South Dakota  
Date Defended: April 1, 2017  

Approved by:
____________________________________  
Major Professor – Kofi R. Bean, Ph.D., Department of Date  
Chemistry and Applied Biological Sciences  
______________________________________________   
Graduate Division Representative – Kaldi T. Goatherd, Date  
Ph.D., Department of Humanities
______________________________________________ 
Committee Member – Skin E. Latte, Ph.D., Department Date
of Civil and Environmental Engineering
______________________________________________ 
Head of the Department of Chemical and Biological Date
Engineering – Dee Caff, Ph.D.
______________________________________________ 
Dean of Graduate Education – Maribeth H. Price, Ph.D. Date
\newpage

# Abstract
<!--Lower case Roman numerals
One page, single-spaced  

The format of the abstract:
Single & separate page, adhering to margin requirements.
Text will be single-spaced.
The Abstract page comes after the Title Page and before the Table of Contents.
An abstract should be informative. It should not state: “this is discussed,” “that was investigated,” or “conclusions are given.” Be as specific on each point as space allows. Avoid stating what the material is about. Concentrate instead on what work has been done. For example, do not write: “the gravity anomalies in the area are discussed.” Rather, state that: “A gravity high of 25 milligals suggests that . . . .”
The most common convention is to report your information impersonally. This, however, is a convention and students should consult with their major professor and graduate committee on possible exceptions. 

(the stuff below is rather old-fashioned, and "I" think they should consider that actions have agency by using first-person voice)
Typically, the restriction on the use of the word “I” and “we” makes it easy to ensure passive voice that is common in abstracts.
Example: “On 30 June a sunspot almost 20 times the size of the planet Earth was observed on the east limb of the sun.” (no clue who did the observation;
vs: On 30 June I observed a sunspot on the east limb of the sun almost 20 times the size of the planet Earth - observer is clearly stated)

The abstract must summarize the following four major elements of the full document:
• Objectives of the research
• Methods used in the research
• Results of the research
• Conclusions and recommendations

The preparation of a good abstract is an art to be acquired through care, thought, and practice. It should convey a maximum amount of information with a minimum number of words, and this requires more writing and re-writing than does almost any other part of the work.
-->
\newpage

# Dedication 
<!--(optional; numbered (Roman))-->
\newpage

# ACKNOWLEDGMENTS 
<!--Lower case Roman numerals
4.14 Acknowledgements
If included, place the acknowledgements between the abstract and the table of contents. Use professional rather than personal writing style to match the rest of the document. Note that externally funded projects generally contractually require an acknowledgement of their financial support.
-->
\newpage

# TABLE OF CONTENTS 
<!--Lower case Roman numerals
The table of contents of this [document](https://www.sdsmt.edu/academics/graduate-education/docs/thesis-and-dissertation-writing-manual/) is an example of proper formatting 
Obviously in the WRONG place!!!
-->
\newpage

# ABBREVIATIONS 
<!--(optional, numbered (Roman))-->
\newpage

# List of Tables
<!--Lower case Roman numerals
-->
You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

<!--Tables should be centered on the page. Captions must also be centered (center justified) and should be placed above the table. Add 1 or 2 empty lines after tables to enable readers to readily differentiate the resumption of narrative text. Oversized tables that are larger than 8 1/2” x 11” (i.e. geographic maps) are to be folded down to a 7” x 10” format.

Tables may be oriented lengthwise on the page, but must be placed in such a manner that when the thesis is held in reading position, the bottom of such a table will be at the right side of the page, the table title is above the table (on the binding side), and the page number is in the same location and angle as for portrait-oriented pages. 
-->
\newpage

# List of Figures
<!--Lower case Roman numerals-->
\newpage




Bibliography that only cites references in the non-journal article chapters, while each journal article must include a self-contained bibliography. 

The student must place a disclaimer at the top of the thesis or dissertation bibliography that says:
“This bibliography is limited to the non-journal article chapters of this thesis/dissertation. 

All journal articles in this thesis/dissertation have a self-contained bibliography.”




\newpage



13.  Bibliography (required, numbered throughout (Arabic))

14.  CURRICULUM VITAE (optional, numbered throughout (Arabic))
4.13 Vita
A vita page, not over one page in length, is to be included as the last page of all theses and dissertations deposited in the Devereaux Library. The vita is to be written in the third person using professional style and could contain the following information (although you may wish to omit A and B if concerned about identity theft):
A. Place and date of birth.
B. Placeanddateofhighschoolgraduation.
C. Place and date of college graduation—with degree and major.
D. Place and date of receipt of master’s degree—with major.
E. Vocational and professional experience (not summer jobs)—including dates, nature of position, and school or organization.
F. Military experience, with indication of professional relevance—if any.
G. Scholarly publications, exhibits of creative work, membership in professional organizations and honorary societies.


Bibliography
\newpage
 
Appendices

Arabic numbers

Label appendix sections with upper case letters (example: “Appendix A: Spectroscopic Data”)
\newpage
Vita
Arabic numbers
\newpage

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

