```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE, message=FALSE}
# Sets up the library of packages 
library("lmomco") # lmoments to find distribution
library("beanplot") # used in lmomco example

library("tidyverse")
library("magrittr")
#library("here") # identifies where to save work
#library("DataExplorer")
#library("rio") # more robust I/O - to import and clean data
#library("janitor") # tools for examining and cleaning dirty data
#library("rnoaa") #  R wrapper for NOAA data inc. NCDC
#library("lubridate")
# library("dataRetrieval") # USGS data import
# library("RColorBrewer") - there is a better one?
# library("workflowr") # creates a research website
# library("colorspace")
# library("bookdown") # 
# library(unpivotr) # fix nasty Excel files
# library("friendlyeval")

# a useful description of commits:
# http://r-pkgs.had.co.nz/git.html

#lmomco <- citation("lmomco")
#toBibtex(lmomco)
```

```{r test-data}
data(amarilloprecip) 
data(canyonprecip) 
data(claudeprecip) 
data(herefordprecip) 
data(tuliaprecip)
data(tulia6Eprecip)
data(vegaprecip)

AMAR <- sort(amarilloprecip$DEPTH)
CANY <- sort(canyonprecip$DEPTH)
CLAU <- sort(claudeprecip$DEPTH)
HERF <- sort(herefordprecip$DEPTH)
TULA <- sort(tuliaprecip$DEPTH)
TUL6 <- sort(tulia6Eprecip$DEPTH)
VEGA <- sort(vegaprecip$DEPTH)

rm(amarilloprecip, canyonprecip, claudeprecip, herefordprecip,
   tuliaprecip, tulia6Eprecip, vegaprecip)

```

```{r test-boxplot}
# combine all into short variable name
x <- list(AMAR = AMAR, CANY = CANY, CLAU = CLAU, HERF = HERF, 
          TULA = TULA, TUL6 = TUL6,VEGA = VEGA) 

# w will be used in a later example
w <- sapply(x,length) 

print(w) 

# show the lengths of the individual records
boxplot(x, ylab="7-DAY ANNUAL MAX RAINFALL, IN INCHES", range=0)
```

```{r tidyverse-data}
# from lmomco book 
data(amarilloprecip) # from lmomco package  
data(canyonprecip)  # .. ditto ..  
data(claudeprecip) # .. ditto .. 
data(herefordprecip) # .. ditto .. 
data(tuliaprecip)  # .. ditto ..  
data(tulia6Eprecip)  # .. ditto ..  
data(vegaprecip) # .. ditto .. 

# put the data together into a single dataframe 
int1 <- full_join(amarilloprecip, canyonprecip, by = "YEAR") %>% 
  rename(AMAR = DEPTH.x) %>% 
  rename(CANY = DEPTH.y) 

int2 <- full_join(int1, claudeprecip, by = "YEAR") %>% 
  rename(CLAU = DEPTH) 

int3 <- full_join(int2, herefordprecip, by = "YEAR") %>% 
  rename(HERF = DEPTH)  

int4 <- full_join(int3, tuliaprecip, by = "YEAR") %>% 
  rename(TULA = DEPTH) 

int5 <- full_join(int4, tulia6Eprecip, by = "YEAR") %>% 
  rename(TUL6 = DEPTH) 

prcp <- full_join(int5, vegaprecip, by = "YEAR") %>%
  rename(VEGA = DEPTH)

rm(amarilloprecip, canyonprecip, claudeprecip, herefordprecip,
   tuliaprecip, tulia6Eprecip, vegaprecip, 
   int1, int2, int3, int4, int5)

# make the wide data long and remove NA vals
prcp_grp <- prcp %>%
  gather(key = "station", value = "depth", -YEAR) %>% 
  drop_na(depth)  

# get the lenths of the datasets
prcp_count <- prcp_grp %>%
  count(station) 
 
# plot the data
ggplot(prcp_grp, aes(as.factor(station), depth)) +
  geom_violin() +
  geom_boxplot() +
  xlab("") +
  ylab("7-day annual max rainfall, inches") +
  theme_bw()
```

```{r plotting_postion}
# find the plotting position: using Weibull; a = 0
prcp_grp <- prcp_grp %>% 
  group_by(station) %>% 
  arrange(depth) %>% 
  mutate(pos = pp(depth)) %>% 
  ungroup() 

# plot the depths as a function of plotting position
ggplot(prcp_grp, aes(pos, depth)) + 
  facet_grid(.~station) + 
  geom_point() + 
  theme_bw() 

```

```{r lmoments}
# This is a long bit of code that takes things out of a list
lmom <- prcp_grp %>%
  split(.$station) %>%
  map(~ lmoms(.$depth)) %>%
  transpose() %>%
  as_tibble() %>%
  select(lambdas, ratios) %>%
    mutate(lambdas = map(lambdas, ~as_tibble(t(.x))))  %>%
    mutate(lambdas = map(lambdas, 
                         ~set_names(.x, c("L1", "L2", "L3", 
                                          "L4", "L5")))) %>%
      mutate(ratios = map(ratios, ~as_tibble(t(.x))))  %>%
    mutate(ratios = map(ratios, 
                         ~set_names(.x, c("T1", "T2", "T3", 
                                          "T4", "T5")))) %>% 
  unnest(lambdas) %>%
  unnest(ratios) %>%
  mutate(station = c("AMAR", "CANY", "CLAU", 
         "HERF", "TULA", "TUL6", "VEGA")) %>%
  select(-T1) %>%
  rename(L_CV = T2) %>% 
  rename(L_skew = T3) %>%
  rename(L_kurtosis = T4) %>%
  select(station, L1, L_CV, L_skew, L_kurtosis)

# Note that we might consider Weiss 1964 value of 1.018

# join the number of years to the station
lmom <- left_join(lmom, prcp_count, by = "station")

rm(prcp_count)

lmrdia <- lmrdia()
```


```{r}

ggplot(lmom, aes(group = station, L_skew, L_kurtosis)) +
  geom_point() +
  theme_classic()
```


```{r test-boxplot, eval=FALSE}
# combine all into short variable name
x <- list(AMAR = AMAR, CANY = CANY, CLAU = CLAU, 
          HERF = HERF,  TULA = TULA, TUL6 = TUL6,VEGA = VEGA) 

# w will be used in a later example
w <- sapply(x,length) 

print(w) 

# show the lengths of the individual records
boxplot(x, ylab="7-DAY ANNUAL MAX RAINFALL, IN INCHES", range=0)
```

```{r bean-plot, eval=FALSE}
rng <- sapply(x, range) # x from previous example
ylim <- c(min(rng[1,]), max(rng[2,]))
par(mfrow=c(2,1), mai=c(0.5,1,0.5,0.5) )
beanplot(x, ll=0.04, main="BEAN PLOT: beanplot()", log="",
 ylim=ylim, ylab="7-DAY ANNUAL MAX RAINFALL,\n IN INCHES",
overallline="median")
```


```{r weibull_lmoment}
AMAR.pp <- pp(AMAR);  CANY.pp <- pp(CANY)
CLAU.pp <- pp(CLAU);  HERF.pp <- pp(HERF)
TULA.pp <- pp(TULA);  TUL6.pp <- pp(TUL6)
VEGA.pp <- pp(VEGA)

AMAR.lmr <- lmoms(AMAR);  CANY.lmr <- lmoms(CANY)
CLAU.lmr <- lmoms(CLAU);  HERF.lmr <- lmoms(HERF)
TULA.lmr <- lmoms(TULA);  TUL6.lmr <- lmoms(TUL6)
VEGA.lmr <- lmoms(VEGA)

```

```{r}
"afunc" <- function(r) {
  return(c(AMAR.lmr$ratios[r], CANY.lmr$ratios[r],
           CLAU.lmr$ratios[r], HERF.lmr$ratios[r],
           TULA.lmr$ratios[r], TUL6.lmr$ratios[r],
           VEGA.lmr$ratios[r]))
}
L1 <- c(AMAR.lmr$lambdas[1], CANY.lmr$lambdas[1],
CLAU.lmr$lambdas[1], HERF.lmr$lambdas[1], TULA.lmr$lambdas[1], TUL6.lmr$lambdas[1], VEGA.lmr$lambdas[1])*1.018; 

print(L1)
# bias correction factor Weiss (1964)
T2 <- afunc(2); T3 <- afunc(3); T4 <- afunc(4)

reg.L1  <- weighted.mean(L1, w);  reg.T2  <- weighted.mean(T2, w)
reg.T3  <- weighted.mean(T3, w);  reg.T4  <- weighted.mean(T4, w)

reg.lmr <- vec2lmom(c(reg.L1, reg.L1*reg.T2, reg.T3, reg.T4)) 
reg.kap <- parkap(reg.lmr) # parameters of the Kappa distribution
str(reg.lmr) # output the regional L-moments List of 9

print(reg.kap) # output the regional Kappa distribution $type
```

```{r}
lmrdia <- lmrdia()
plotlmrdia(lmrdia,autolegend=TRUE, nopoints=TRUE,
           nolimits=TRUE, xlim=c(0,0.3), ylim=c(-0.1,0.4),
           xleg=0.05, yleg=0.3)
points(T3,T4)
points(reg.T3,reg.T4, pch=16, cex=2)
plotlmrdia(lmrdia()) # simplest of all uses
```

## Overview
<!--This is the chapter 2 hydrologic drought.  
Next Steps:
1. Methods.
2. Find the ramblings I wrote about scope and purpose
3. figure out how to reference stuff 
-->

## Introduction


## Methods
I imported Global Historical Climatology Network (GHCN) daily precipitation records for candidate "WEATHER STATIONS" into R-Studio (REF1) using the "rnoaa" package.
I used Theissen polygons and the length and continuity of precipitation records to select stations for further analysis.
I used 'dplyr' to fill NA values with data from nearest station
I used 'dplyr' to create monthly vals from daily vals.
I removed short records: Oral & Long Valley after checking for 
covariance.

<!-- 
Work on finishing describing methods for EDA.
-->

## Results
<!--
Work on describing results of EDA.
