```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE, message=FALSE}
# Sets up the library of packages 
library("lmomco") # lmoments to find distribution
library("tidyverse")
library("magrittr") # contains easier ways to say things about lists
library("here") # identifies where to save work
library("rio") # more robust I/O - to import and clean data
library("lubridate")

#lmomco <- citation("lmomco")
#toBibtex(lmomco)
```

<!--
Exploratory Data Analysis Checklist by Roger Peng 
https://leanpub.com/exdata  

1.0  Formulate your question  
Spend a few minutes to figure out the question you’re really interested in, and narrow it down to be as specific as possible (without becoming uninteresting).

General question:
Are air pollution levels higher on the east coast than on the west coast?
More specific question:
Are hourly ozone levels on average higher in New York City than they are in Los Angeles?

2.0   Read in your data  
Sometimes the data will need some cleaning and every dataset has its unique quirks. The dataset is a comma-separated value (CSV) file, where each row of the file contains one daily measurement of precipation depth.  The readr package can rewrite column names to remove spaces.
> names(ozone) <- make.names(names(ozone))

3.0  Check the dataset 
3.1  Check the number of rows and columns.
3.2  Check the types of data
3.3  Look at the top and the bottom of your data 
3.4  Check your “n”s & NAs 
3.5  Validate with at least one external data source  
4.0  Try the easy solution first to answer question
5.0  Challenge your solution 
6.0  Follow up questions 
-->

## Broad questions:
What is the drought history of the Pine Ridge Reservation?  
Does the drought extent differ across the study area?

## Narrower question:
What is the underlying distribution of precipitation data?  

## Analysis Steps:
1. Recreated analysis from the lmomco text ch 12 (author?) in Tidyverse
2. Imported cleaned precipitation data (see 04_prcp-data_munging)   
3. Applied a sqrt transform to reduce skew
4. 
5. 
6. 

## Variable naming convention:   
sta          precipitation station  
_meta        metadata  
_mon         monthly precipitation depths  
_grp         wide data changed to long data
_notzero     non-zero precip values
_zero        precip values equal to zero
_log         log10 of monthly precipitation depths
_count       number of months in a given record

lmom_sta     L-moments for stations
lmom_reg     Weighted mean of L-moments for stations

min          minimum non-zero value
L1           first L-moment, similiar to mean
L_CV         first L-moment ratio, similiar to coefficient of var
L_skew       second L-moment ratio, similiar to skewness 
L_kurtosis   third L-moment ratio, similiar to kurtosis
n            number of months in a given record
int#         intermediate variable used to bind rows; # = 1, 2, ...

# Used in the L-moment diagram
aep4         4-Parameter Asymmetric Exponential Power Distribution 
gev          Generalized Extreme Value Distribution
glo          Generalized Logistic Distribution
gpa          Generalized Pareto Distribution
####gno      Generalized Logistic Distribution
gov          Govindarajulu Distribution
pe3          Pearson Type III Distribution


## Thoughts - the orig depth vs plotting vals look j-shaped.  
## Sqrt transform vs plotting vals looks sinusoidal
## Log-tranfrom
## Results: 


```{r import-data}
# General Purpose: prepare data for drought index  
# Specific purpose: graphical EDA  
sta_meta    <- as.tibble(import("data/sta_meta_fin3.csv"))  
sta_mon     <- as.tibble(import("data/stations_monthly.csv")) 

# fix date & add year and month  
sta_mon <- sta_mon %>% 
  mutate(date = ymd(date)) %>% 
  arrange(desc(date)) 

# make the wide data long, remove NA vals, sqrt transform
sta_grp <- sta_mon %>%
  gather(key = "station", value = "depth", -date, -year, -month) %>%
  drop_na(depth) %>% 
  mutate(sqrt_depth = sqrt(depth)) 

# Check on log transformation
sta_notzero <- sta_grp %>%
  filter(depth != 0)

min <- min(sta_notzero$depth)

sta_zero <- sta_grp %>%
  filter(depth == 0) %>%
  mutate(depth = depth + min/2)

sta_log <- bind_rows(sta_zero, sta_notzero)
sta_log <- sta_log %>%
  mutate(log_depth = log10(depth))

rm(min)

# plot the data
ggplot(sta_log, aes(as.factor(station), log_depth)) +
  geom_violin() +
  geom_boxplot() +
  xlab("") +
#  ylab("Sqrt of monthly rainfall depths, originally in mm") +
#  scale_y_sqrt() +
  scale_y_log10() +
  theme_bw()
```

```{r station_plotting_postion}
# find the plotting position: using Weibull; a = 0
sta_grp <- sta_grp %>% 
  group_by(station) %>% 
  arrange(log_depth) %>% 
  mutate(position = pp(log_depth)) %>% 
  ungroup() 

# find the plotting position: using Weibull; a = 0
sta_log <- sta_log %>% 
  group_by(station) %>% 
  arrange(log_depth) %>% 
  mutate(position = pp(log_depth)) %>% 
  ungroup() 


# plot the depths as a function of plotting position
ggplot(sta_log, aes(position, log_depth)) + 
  facet_grid(.~station) + 
  geom_point() + 
  theme_bw() 
```

```{r lmoments}
# This is a long bit of code that takes things out of a list

lmom_sta <- sta_log %>%
  split(.$station) %>%
  map(~ lmoms(.$log_depth)) %>%
  transpose() %>%
  as_tibble() %>%
  select(lambdas, ratios) %>%
    mutate(lambdas = map(lambdas, ~as_tibble(t(.x))))  %>%
    mutate(lambdas = map(lambdas, 
                         ~set_names(.x, c("L1", "L2", "L3", 
                                          "L4", "L5")))) %>%
      mutate(ratios = map(ratios, ~as_tibble(t(.x))))  %>%
    mutate(ratios = map(ratios, 
                         ~set_names(.x, c("T1", "T2", "T3", 
                                          "T4", "T5")))) %>% 
  unnest(lambdas) %>%
  unnest(ratios) %>%
  mutate(station = c("oel", "cot", "rap", 
         "int", "ora")) %>%
  select(-T1) %>%
  rename(L_CV = T2) %>% 
  rename(L_skew = T3) %>%
  rename(L_kurtosis = T4) %>%
  select(station, L1, L_CV, L_skew, L_kurtosis)

# Note that we might consider Weiss 1964 bias value of 1.018 for L1

# get the lenths of the datasets
sta_count <- sta_log %>%
  count(station)  

# join the number of years to the station
lmom_sta <- left_join(lmom_sta, sta_count, by = "station")
rm(prcp_count)

# calculate weighted means
L1  <- weighted.mean(lmom_sta$L1, lmom_sta$n)
L_CV  <- weighted.mean(lmom_sta$L_CV, lmom_sta$n)
L_skew  <- weighted.mean(lmom_sta$L_skew, lmom_sta$n)
L_kurtosis  <- weighted.mean(lmom_sta$L_kurtosis, lmom_sta$n)
n       <- sum(lmom_sta$n)

# combine
int1     <- cbind(L1, L_CV)
int2     <- cbind(int1, L_skew)
int3     <- cbind(int2, L_kurtosis)
lmom_reg <- cbind(int3, n)

rm(L1, L_CV, L_skew, L_kurtosis, n, int1, int2, int3)

lmom_reg <- as.tibble(lmom_reg) %>%
  mutate(station = "WtMean") %>%
  select(station, everything())

# lmom_sta <- bind_rows(lmom_sta, lmom_reg)
```

```{r elmoment_diagram_ratios}
# extract elements from the lmrdia list to plot in ggplot2 
lmrdia <- lmrdia()

aep4 <- lmrdia %>%
  extract2(2) %>%
  as.tibble()

gev <- lmrdia %>%
  extract2(5) %>%
  as.tibble()

glo <- lmrdia %>%
  extract2(6) %>%
  as.tibble()

gpa <- lmrdia %>%
  extract2(7) %>%
  as.tibble()

gno <- lmrdia %>%
  extract2(9) %>%
  as.tibble()

gov <- lmrdia %>%
  extract2(10) %>%
  as.tibble()

pe3 <- lmrdia %>%
  extract2(12) %>%
  as.tibble()
```

```{r plot-lmoment-diagram}

ggplot() +
  geom_line(data = aep4, aes(V1, V2, color = "red")) + 
  geom_line(data = gev, aes(V1, V2, color = "blue")) + 
  geom_line(data = glo, aes(V1, V2, color = "orange")) + 
  geom_line(data = gpa, aes(V1, V2, color = "violet")) + 
  geom_line(data = gno, aes(V1, V2, color = "green")) + 
  geom_line(data = gov, aes(V1, V2, color = "brown")) + 
  geom_line(data = pe3, aes(V1, V2)) + 
  geom_point(data = lmom_sta, aes(L_skew, L_kurtosis)) +
  geom_point(data = lmom_reg, aes(L_skew, L_kurtosis, size = 3)) +
  theme_bw() +
  xlim(-0.3, 0) +
  ylim(0, 0.3)

# ggsave()
```

```{r tidyverse-data}
# from lmomco book 
data(amarilloprecip) # from lmomco package  
data(canyonprecip)  # .. ditto ..  
data(claudeprecip) # .. ditto .. 
data(herefordprecip) # .. ditto .. 
data(tuliaprecip)  # .. ditto ..  
data(tulia6Eprecip)  # .. ditto ..  
data(vegaprecip) # .. ditto .. 

# put the data together into a single dataframe 
int1 <- full_join(amarilloprecip, canyonprecip, by = "YEAR") %>% 
  rename(AMAR = DEPTH.x) %>% 
  rename(CANY = DEPTH.y) 

int2 <- full_join(int1, claudeprecip, by = "YEAR") %>% 
  rename(CLAU = DEPTH) 

int3 <- full_join(int2, herefordprecip, by = "YEAR") %>% 
  rename(HERF = DEPTH)  

int4 <- full_join(int3, tuliaprecip, by = "YEAR") %>% 
  rename(TULA = DEPTH) 

int5 <- full_join(int4, tulia6Eprecip, by = "YEAR") %>% 
  rename(TUL6 = DEPTH) 

prcp <- full_join(int5, vegaprecip, by = "YEAR") %>%
  rename(VEGA = DEPTH)

rm(amarilloprecip, canyonprecip, claudeprecip, herefordprecip,
   tuliaprecip, tulia6Eprecip, vegaprecip, 
   int1, int2, int3, int4, int5)

# make the wide data long and remove NA vals
prcp_grp <- prcp %>%
  gather(key = "station", value = "depth", -YEAR) %>% 
  drop_na(depth)  
 
# plot the data
ggplot(prcp_grp, aes(as.factor(station), depth)) +
  geom_violin() +
  geom_boxplot() +
  xlab("") +
  ylab("7-day annual max rainfall, inches") +
  theme_bw()
```

```{r plotting_postion}
# find the plotting position: using Weibull; a = 0
prcp_grp <- prcp_grp %>% 
  group_by(station) %>% 
  arrange(depth) %>% 
  mutate(pos = pp(depth)) %>% 
  ungroup() 

# plot the depths as a function of plotting position
ggplot(prcp_grp, aes(pos, depth)) + 
  facet_grid(.~station) + 
  geom_point() + 
  theme_bw() 
```

```{r lmoments}
# This is a long bit of code that takes things out of a list
lmom <- prcp_grp %>%
  split(.$station) %>%
  map(~ lmoms(.$depth)) %>%
  transpose() %>%
  as_tibble() %>%
  select(lambdas, ratios) %>%
    mutate(lambdas = map(lambdas, ~as_tibble(t(.x))))  %>%
    mutate(lambdas = map(lambdas, 
                         ~set_names(.x, c("L1", "L2", "L3", 
                                          "L4", "L5")))) %>%
      mutate(ratios = map(ratios, ~as_tibble(t(.x))))  %>%
    mutate(ratios = map(ratios, 
                         ~set_names(.x, c("T1", "T2", "T3", 
                                          "T4", "T5")))) %>% 
  unnest(lambdas) %>%
  unnest(ratios) %>%
  mutate(station = c("AMAR", "CANY", "CLAU", 
         "HERF", "TULA", "TUL6", "VEGA")) %>%
  select(-T1) %>%
  rename(L_CV = T2) %>% 
  rename(L_skew = T3) %>%
  rename(L_kurtosis = T4) %>%
  select(station, L1, L_CV, L_skew, L_kurtosis)

# Note that we might consider Weiss 1964 bias value of 1.018 for L1

# get the lenths of the datasets
prcp_count <- prcp_grp %>%
  count(station)  

# join the number of years to the station
lmom <- left_join(lmom, prcp_count, by = "station")
rm(prcp_count)

# calculate weighted means
L1  <- weighted.mean(lmom$L1, lmom$n)
L_CV  <- weighted.mean(lmom$L_CV, lmom$n)
L_skew  <- weighted.mean(lmom$L_skew, lmom$n)
L_kurtosis  <- weighted.mean(lmom$L_kurtosis, lmom$n)
n       <- sum(lmom$n)

# combine
int1     <- cbind(L1, L_CV)
int2     <- cbind(int1, L_skew)
int3     <- cbind(int2, L_kurtosis)
lmom_reg <- cbind(int3, n)

rm(L1, L_CV, L_skew, L_kurtosis, n, int1, int2, int3)

lmom_reg <- as.tibble(lmom_reg) %>%
  mutate(station = "WtMean") %>%
  select(station, everything())

lmom <- bind_rows(lmom, lmom_reg)
```

```{r extract_lmoment_diagram_ratios}
# extract elements from the lmrdia list to plot in ggplot2 
lmrdia <- lmrdia()

aep4 <- lmrdia %>%
  extract2(2) %>%
  as.tibble()

gev <- lmrdia %>%
  extract2(5) %>%
  as.tibble()

glo <- lmrdia %>%
  extract2(6) %>%
  as.tibble()

gpa <- lmrdia %>%
  extract2(7) %>%
  as.tibble()

gno <- lmrdia %>%
  extract2(9) %>%
  as.tibble()

gov <- lmrdia %>%
  extract2(10) %>%
  as.tibble()

pe3 <- lmrdia %>%
  extract2(12) %>%
  as.tibble()
```

```{r plot-lmoment-diagram}

ggplot() +
  geom_line(data = aep4, aes(V1, V2, color = "red")) + 
  geom_line(data = gev, aes(V1, V2, color = "blue")) + 
  geom_line(data = glo, aes(V1, V2, color = "orange")) + 
  geom_line(data = gpa, aes(V1, V2, color = "violet")) + 
  geom_line(data = gno, aes(V1, V2, color = "green")) + 
  geom_line(data = gov, aes(V1, V2, color = "brown")) + 
  geom_line(data = pe3, aes(V1, V2)) + 
  geom_point(data = lmom, aes(L_skew, L_kurtosis)) +
  geom_point(data = lmom_reg, aes(L_skew, L_kurtosis, size = 3)) +
  theme_bw() +
  xlim(0, 0.3) +
  ylim(-0.1, 0.4)
```


```{r test-boxplot, eval=FALSE}
# combine all into short variable name
x <- list(AMAR = AMAR, CANY = CANY, CLAU = CLAU, 
          HERF = HERF,  TULA = TULA, TUL6 = TUL6,VEGA = VEGA) 

# w will be used in a later example
w <- sapply(x,length) 

print(w) 

# show the lengths of the individual records
boxplot(x, ylab="7-DAY ANNUAL MAX RAINFALL, IN INCHES", range=0)
```

```{r bean-plot, eval=FALSE}
rng <- sapply(x, range) # x from previous example
ylim <- c(min(rng[1,]), max(rng[2,]))
par(mfrow=c(2,1), mai=c(0.5,1,0.5,0.5) )
beanplot(x, ll=0.04, main="BEAN PLOT: beanplot()", log="",
 ylim=ylim, ylab="7-DAY ANNUAL MAX RAINFALL,\n IN INCHES",
overallline="median")
```


```{r weibull_lmoment}
AMAR.pp <- pp(AMAR);  CANY.pp <- pp(CANY)
CLAU.pp <- pp(CLAU);  HERF.pp <- pp(HERF)
TULA.pp <- pp(TULA);  TUL6.pp <- pp(TUL6)
VEGA.pp <- pp(VEGA)

AMAR.lmr <- lmoms(AMAR);  CANY.lmr <- lmoms(CANY)
CLAU.lmr <- lmoms(CLAU);  HERF.lmr <- lmoms(HERF)
TULA.lmr <- lmoms(TULA);  TUL6.lmr <- lmoms(TUL6)
VEGA.lmr <- lmoms(VEGA)

```

```{r}
"afunc" <- function(r) {
  return(c(AMAR.lmr$ratios[r], CANY.lmr$ratios[r],
           CLAU.lmr$ratios[r], HERF.lmr$ratios[r],
           TULA.lmr$ratios[r], TUL6.lmr$ratios[r],
           VEGA.lmr$ratios[r]))
}
L1 <- c(AMAR.lmr$lambdas[1], CANY.lmr$lambdas[1],
CLAU.lmr$lambdas[1], HERF.lmr$lambdas[1], TULA.lmr$lambdas[1], TUL6.lmr$lambdas[1], VEGA.lmr$lambdas[1])*1.018; 

print(L1)
# bias correction factor Weiss (1964)
T2 <- afunc(2); T3 <- afunc(3); T4 <- afunc(4)

reg.L1  <- weighted.mean(L1, w);  reg.T2  <- weighted.mean(T2, w)
reg.T3  <- weighted.mean(T3, w);  reg.T4  <- weighted.mean(T4, w)

reg.lmr <- vec2lmom(c(reg.L1, reg.L1*reg.T2, reg.T3, reg.T4)) 
reg.kap <- parkap(reg.lmr) # parameters of the Kappa distribution
str(reg.lmr) # output the regional L-moments List of 9

print(reg.kap) # output the regional Kappa distribution $type
```

```{r}
lmrdia <- lmrdia()
plotlmrdia(lmrdia,autolegend=TRUE, nopoints=TRUE,
           nolimits=TRUE, xlim=c(0,0.3), ylim=c(-0.1,0.4),
           xleg=0.05, yleg=0.3)
points(T3,T4)
points(reg.T3,reg.T4, pch=16, cex=2)
plotlmrdia(lmrdia()) # simplest of all uses
```

## Overview
<!--This is the chapter 2 hydrologic drought.  
Next Steps:
1. Methods.
2. Find the ramblings I wrote about scope and purpose
3. figure out how to reference stuff 
-->

## Introduction


## Methods
I imported Global Historical Climatology Network (GHCN) daily precipitation records for candidate "WEATHER STATIONS" into R-Studio (REF1) using the "rnoaa" package.
I used Theissen polygons and the length and continuity of precipitation records to select stations for further analysis.
I used 'dplyr' to fill NA values with data from nearest station
I used 'dplyr' to create monthly vals from daily vals.
I removed short records: Oral & Long Valley after checking for 
covariance.

<!-- 
Work on finishing describing methods for EDA.
-->

## Results
<!--
Work on describing results of EDA.
