<!--
Exploratory Data Analysis Checklist by Roger Peng 
https://leanpub.com/exdata  

1.0  Formulate your question  
2.0   Read in your data  
3.0  Check the dataset 
3.1  Check the number of rows and columns.
3.2  Check the types of data
3.3  Look at the top and the bottom of your data 
3.4  Check your “n”s & NAs 
3.5  Validate with at least one external data source  
4.0  Try the easy solution first to answer question
5.0  Challenge your solution 
6.0  Follow up questions 

## Broad questions:
What is the drought history of the Pine Ridge Reservation?  
Does the drought extent differ across the study area?

## Narrower question:
What is the underlying distribution of precipitation data?  

## Analysis Steps:
1. Recreated analysis from the lmomco text ch 12 (author?) in Tidyverse
2. Imported cleaned precipitation data (see 04_prcp-data_munging)   
3. Applied sqrt & log10 transform to explore effects on skew 
4. Explored the data with box plots, violin plot.
5. Applied Weibull plotting position and graphed the data on sqrt plot
6. Calculated L-moments and L-moment ratios
7. Plotted the 

## Variable naming convention:   
sta          precipitation station  
_meta        metadata  
_mon         monthly precipitation depths  
_grp         wide data changed to long data
_notzero     non-zero precip values
_zero        precip values equal to zero
_log         log10 of monthly precipitation depths
_count       number of months in a given record

lmom_sta     L-moments for stations
lmom_reg     Weighted mean of L-moments for stations

min          minimum non-zero value
L1           first L-moment, similiar to mean
L_CV         first L-moment ratio, similiar to coefficient of var
L_skew       second L-moment ratio, similiar to skewness 
L_kurtosis   third L-moment ratio, similiar to kurtosis
n            number of months in a given record
int#         intermediate variable used to bind rows; # = 1, 2, ...

# Used in the L-moment diagram
aep4         4-Parameter Asymmetric Exponential Power Distribution 
gev          Generalized Extreme Value Distribution
glo          Generalized Logistic Distribution
gpa          Generalized Pareto Distribution
####gno      Generalized Logistic Distribution #need to fix as GLO
gov          Govindarajulu Distribution
pe3          Pearson Type III Distribution


## Thoughts - the orig depth vs plotting vals look j-shaped.  
## Sqrt trans vs plotting vals look slightly sinusoidal; nice boxplots
## Log-tranformation looks like the mirror of orig depth vs plotting

## Results: Fits a PE3 distribution
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r library, include=FALSE, message=FALSE}
# Sets up the library of packages 
library("lmomco") # lmoments to find distribution
library("tidyverse")
library("magrittr") # contains easier ways to say things about lists
library("here") # identifies where to save work
library("rio") # more robust I/O - to import and clean data
library("lubridate")

#lmomco <- citation("lmomco")
#toBibtex(lmomco)
```

```{r import-data}
# General Purpose: prepare data for drought index  
# Specific purpose: graphical EDA  
sta_meta    <- as.tibble(import("data/sta_meta_fin3.csv"))  
sta_mon     <- as.tibble(import("data/stations_monthly.csv")) 

# fix date & add year and month  
sta_mon <- sta_mon %>% 
  mutate(date = ymd(date)) %>% 
  arrange(desc(date)) 

# make the wide data long, remove NA vals, sqrt transform
sta_grp <- sta_mon %>%
  gather(key = "station", value = "depth", -date, -year, -month) %>%
  drop_na(depth)  # %>% 
#  mutate(sqrt_depth = sqrt(depth)) 

# Check on log transformation
#sta_notzero <- sta_grp %>%
#  filter(depth != 0)

#min <- min(sta_notzero$depth)

#sta_zero <- sta_grp %>%
#  filter(depth == 0) %>%
#  mutate(depth = depth + min/2)

#sta_log <- bind_rows(sta_zero, sta_notzero)
#sta_log <- sta_log %>%
#  mutate(log_depth = log10(depth))

#rm(min)

# plot the data
ggplot(sta_grp, aes(as.factor(station), depth)) +
  geom_violin() +
  geom_boxplot() +
#  scale_y_sqrt() +
#  scale_y_log10() +
  theme_bw() +
  ggtitle("Weather stations near Pine Ridge Reservation, SD", 
          subtitle = "1909-2018") +
  xlab("") +
  ylab("Monthly depth in mm")
```

```{r station_plotting_postion}
# find the plotting position: using Weibull; a = 0
sta_grp <- sta_grp %>% 
  group_by(station) %>% 
  arrange(depth) %>% 
  mutate(position = pp(depth)) %>% 
  ungroup() 

# find the plotting position for log-data: using Weibull; a = 0
#sta_log <- sta_log %>% 
#  group_by(station) %>% 
# arrange(log_depth) %>% 
#  mutate(position = pp(log_depth)) %>% 
#  ungroup() 

# plot the depths as a function of plotting position
ggplot(sta_grp, aes(position, depth)) + 
  facet_grid(.~station) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_bw() +
ggtitle("Weather stations near Pine Ridge Reservation, SD", 
          subtitle = "1909-2018") + 
  xlab("Frequency of occurrance") +
  ylab("Monthly depth in mm") +
  scale_y_sqrt() +
  ylab("Depth in mm")
```

```{r lmoments}
# This is a long bit of code that takes things out of a list

lmom_sta <- sta_grp %>%
  split(.$station) %>%
  map(~ lmoms(.$depth)) %>%
  transpose() %>%
  as_tibble() %>%
  select(lambdas, ratios) %>%
    mutate(lambdas = map(lambdas, ~as_tibble(t(.x))))  %>%
    mutate(lambdas = map(lambdas, 
                         ~set_names(.x, c("L1", "L2", "L3", 
                                          "L4", "L5")))) %>%
      mutate(ratios = map(ratios, ~as_tibble(t(.x))))  %>%
    mutate(ratios = map(ratios, 
                         ~set_names(.x, c("T1", "T2", "T3", 
                                          "T4", "T5")))) %>% 
  unnest(lambdas) %>%
  unnest(ratios) %>%
  mutate(station = c("oel", "cot", "rap", 
         "int", "ora")) %>%
  select(-T1) %>%
  rename(L_CV = T2) %>% 
  rename(L_skew = T3) %>%
  rename(L_kurtosis = T4) %>%
  select(station, L1, L_CV, L_skew, L_kurtosis)

# Note that we might consider Weiss 1964 bias value of 1.018 for L1

# get the lenths of the datasets
sta_count <- sta_grp %>%
  count(station)  

# join the number of years to the station
lmom_sta <- left_join(lmom_sta, sta_count, by = "station")
rm(prcp_count)

# calculate weighted means for regional L-moments 
L1  <- weighted.mean(lmom_sta$L1, lmom_sta$n)
L_CV  <- weighted.mean(lmom_sta$L_CV, lmom_sta$n)
L_skew  <- weighted.mean(lmom_sta$L_skew, lmom_sta$n)
L_kurtosis  <- weighted.mean(lmom_sta$L_kurtosis, lmom_sta$n)
n       <- sum(lmom_sta$n)

# combine the output into a single weighted mean
int1     <- cbind(L1, L_CV)
int2     <- cbind(int1, L_skew)
int3     <- cbind(int2, L_kurtosis)
lmom_reg <- cbind(int3, n)

rm(L1, L_CV, L_skew, L_kurtosis, n, int1, int2, int3, sta_count)

# finalize the regional L-moment
lmom_reg <- as.tibble(lmom_reg) %>%
  mutate(station = "WtMean") %>%
  select(station, everything())
```

```{r Lmoment_diagram_ratios}
# extract elements from the lmrdia list to plot in ggplot2 
#   the x-value is the L-skewness and y-value is L-kurtosis 

# get vals from the lmrdia list
lmrdia <- lmrdia() 

# extract L-skew & L-kurtosis values for several distributions
aep4 <- lmrdia %>%
  extract2(2) %>%
  as.tibble()

gev <- lmrdia %>%
  extract2(5) %>% as.tibble()

glo <- lmrdia %>%
  extract2(6) %>% as.tibble()

gpa <- lmrdia %>%
  extract2(7) %>% as.tibble()

gno <- lmrdia %>%
  extract2(9) %>% as.tibble()

gov <- lmrdia %>%
  extract2(10) %>% as.tibble()

pe3 <- lmrdia %>%
  extract2(12) %>% as.tibble()

# combine and rename columns as distribution types
int1      <- full_join(aep4, gev, by = "V1")  
int1      <- int1 %>% 
               rename(AEP4 = V2.x) %>% 
               rename(GEV = V2.y) 

int2      <- full_join(int1, glo, by = "V1") 
int2      <- int2 %>% rename(GLO = V2) 

int3      <- full_join(int2, gpa, by = "V1") 
int3      <- int3 %>% rename(GPA = V2) 

int4      <- full_join(int3, gno, by = "V1") 
int4      <- int4 %>% rename(GNO = V2) 

int5      <- full_join(int4, gov, by = "V1") 
int5      <- int5 %>% rename(GOV = V2) 

lmom_theo <- full_join(int5, pe3, by = "V1")
lmom_theo <- lmom_theo %>% rename(PE3 = V2) %>% 
  rename(L_skew = V1) %>% 
  arrange(L_skew)

# prepare theoretical distributions for plotting
lmom_theo <- lmom_theo %>%
  gather(key = "distribution", value = "L_kurtosis", -L_skew) %>%
  drop_na(L_kurtosis) %>%
  select(distribution, everything()) 

rm(aep4, gev, int1, glo, int2, gpa,int3, gno, int4, gov, 
   int5, pe3, lmrdia)
```

```{r plot-lmoment-diagram}
# plots the theo distributions, the sample vals, and regional mean
ggplot() +
  geom_line(data = lmom_theo, aes(L_skew, L_kurtosis, 
                                  group = distribution, 
                                  linetype = distribution)) +
  geom_point(data = lmom_sta, aes(L_skew, L_kurtosis)) +
  geom_point(data = lmom_reg, aes(L_skew, L_kurtosis, 
                                  size = 3, show.legend = NA)) +
  theme_bw() +
  xlim(0.25, 0.5) +
  ylim(0, 0.25) +
  ggtitle("L-moment diagram for monthly precipitation depth", 
          subtitle = "Weather stations near Pine Ridge Reservation, 1909-2018")

```

## Overview
<!--This is the chapter 2 hydrologic drought.  
Next Steps:
1. Methods.
2. Find the ramblings I wrote about scope and purpose
3. figure out how to reference stuff 
-->

## Introduction


## Methods
I imported Global Historical Climatology Network (GHCN) daily precipitation records for candidate "WEATHER STATIONS" into R-Studio (REF1) using the "rnoaa" package.
I used Theissen polygons and the length and continuity of precipitation records to select stations for further analysis.
I used 'dplyr' to fill NA values with data from nearest station
I used 'dplyr' to create monthly vals from daily vals.
I removed short records: Oral & Long Valley after checking for 
covariance.

<!-- 
Work on finishing describing methods for EDA.
-->

## Results
<!--
Work on describing results of EDA.
