---
title: "02-chapt2-5_study-area"
author: "CJ Tinant"
date: "8/06/2019"
output: html_document 
---

<!--
Purpose: 
1) Develop tables and plots for the study area description section of Chapter 2 
2) Refactor prior code 

Data:
Predominant datasets used are:  
1) USGS daily streamflow and station metadata,  
2) Summary data from a QGIS analysis of ungaged watersheds of interest.  

Approach: 
1) find potential USGS gages by bounding box 
     bBox = a contiguous range of decimal latitude and longitude, starting 
     with the west longitude, then the south latitude, then the east longitude, 
     and then the north latitude with each value separated by a comma. 
# https://waterservices.usgs.gov/rest/Site-Service.html#bBox 
# the Pine Ridge Reservation boundary is c(-103.0, 43.0, -100.2, 43.8)

2) filter gages for the following reasons
     - upstream control, 
     - in crystaline or karst catchments
     - provisional data 
     - (sigh) overlooked in the first data screening from 1990-2017 

3) clean metadata by removing non-needed variables  

4) obtain USGS gage daily flow data

5) filter stations based on record length 
    - no data gages for 1980-2018 water years 
    - less than 6-years of continuous data for 1990-2018
      for filtering info see: https://help.waterdata.usgs.gov/site_tp_cd 

6) create metadata tables 
    - gage_table - summary of gage metadata 
    - ung_table - summary of ungaged watershed metadata 
    - soils_table - summary of geology, soils, and vegetation 

7) prepare daily flow data for PCA 
     - check missing records & filter 
     - remove provincial data 
     - eliminate watershed size effects: 
          Daily flow depth is calculated by dividing flow (cms) by watershed 
          area (sq-km) and multiply the resultant by num. of sec. in a day.  
          The resulting unit is cu-m-d per sq-km. 
     - transform the data by BoxCox transformation by forcast::BoxCox
          The daily flow data are highly skewed.  BoxCox transformation uses 
          maximum likelihood to identify a best value (lambda) that best 
          transforms a dataset to an ~ normal distribution.

          # this is table of what the lambda values (probably) mean
          # lambda = 1 is normal distribution (no change), 
          # lambda = 0.5 is a square-root transformation, 
          # lamda = 2 is a square transformation,
          # lambda = 0 is a logrithmic transformation.

8) make a PCA matrix and tidy the output 
     - used prcomp() with center & scale = TRUE to center the data with a 
         mean of zero and standard deviation of unity, e.g. a z-score.  
         Note: '.' passes select(q1_tr, q7_tr, q30_tr) %>% to prcomp() 
     - found eigenvectors-- the results about the PC axes

Next Steps: 
-- check into stylr-packag

Variable naming convention:   
# ~~~~~~~~~~~~~~~~~~~~~~~~~
gage_poss           possible USGS gaging stations in the study area 
gage_meta_poss      metadata for possible gaging stations
-- _int1            scratch df for pulling gages with integer fields 
-- _int2            scratch df for pulling gages with integer fields 
-- _char            scratch df for pulling gages with character fields 

gage_meta           cleaned metadata for USGS gaging stations in the study area 
  "site_no"                site number     
  "station_nm"             station name     
  "dec_lat_va"             latitude value in decimal degrees 
  "dec_long_va"            longitude value in decimal degrees 
  "state_cd"               State code
  "county_cd"              County code
  "alt_va"                 altitude 
  "huc_cd"                 hydrologic unit code 
  "drain_area_va"          drainage area ???in square miles???
  "contrib_drain_area_va"  contributing drainage area ???in square miles???

endDate             used for calling Egret::readNWISDaily 
parameter_cd        used for calling Egret::readNWISDaily 
startDate           used for calling Egret::readNWISDaily 

lon_riv_dv          scratch df for pulling lon_riv 
gage_most_dv        scratch df for pulling !lon_riv 
yrs_rec             calculates years of record 
gage_dv             daily flow values for waterYear 1980-2017 

wsd_summary     zonal statistics of watershed environmental parameters 
*   "_id"       unique id
*  "_sta_id"    four- or seven-digit station ID   
   "_type "     distinguishes gaged from ungaged 
   "_watshed"   describes the HUC06 watershed 
   "_HUC12"     hydrologic unit code 12
   "_sta_name"    station name derived from outlet HUC12 catchment name
   "_gage_num"    USGS site number 
   "_gage_nm"     station name from USGS gage
   "_dec_lat"     latitude in decimal degrees
   "_dec_lon"     longitude in decimal degrees
   "_cat_area"    catchment area in sq km
*  "_cat_area_l"  natural logarithm of catchment area 
   "_cat_length"  catchment length
   "_cat_width"   catchment width
*  "_lw_ratio"    catchment length divided by catchment width
   "_str_len"     stream length
*  "_drain_dens"  stream length divided by catchment area
*  "_prcp_mean"   areal mean precipitation depth 1992-2012
*  "_t07_mean"    average July temperature 1992-2012
*  "_vpd_ann"     areal mean of max vapor deficit 1992-2012
*  _"vpd_07"      areal mean of max July vapor deficit 1992-2012
*  "_cat_out"     catchment outlet elevation
*  "_cat_rel"     difference in max and min elevation
*  "_slop_med"    median percent slope
*  "_TWI_mean"    mean terrain wetness index
*  "_perc_cov"    percent forest cover from NLCD 2016
*  "_fc_mean"     mean field capacity
*  "_ksat_mean"   mean horizontal saturated hydraulic conductivity
*  "_kvert_mean"  mean vertical saturated hydrologic conductivity 

gage_table        summary of gage metadata for flextable 
ung_table         summary of ungaged watershed metadata for flextable  
soils_table       summary of geology, soils, and vegetation for flextable  

the tables above use flextable control keys: 
col_key_char      controls character elelements for flextables
col_key_int       controls integer elelements for flextables
col_key_num       controls numeric elelements for flextables

gage_check        scratch variable used to check operations 
gage_raw          the "raw" data, after removing the NA Qvals 
  _site_no        Eight-digit USGS gage number
  _sta            Short name used for this project 
  _Date           Date 
  _Q              Discharge in m3/s
  _Julian         Num. days since January 1, 1850 
  _Month          Month of the year [1-12] 
  _Day            Day of the year [1-366] 
  _DecYear        Decimal year 
  _MonthSeq       Number of months since January 1, 1850 
  _waterYear      Year following USGS water year 
  _Qualifier      Qualifying code
  _ i             Index of days, starting with 1 
  _LogQ           Natural logarithm of Q 
  _Q7             7-day running average of Q
  _Q30            30-day running average of Q

gage_dv_gath      scratch variable, gathers the Q, Q7, Q30 
gage_dv_low       censored zero flows (q_val < 0.01)
gage_dv_high      filtered non-zero flows -> join flows 
lambda_qval       BoxCox lambda value for qval 
gage_contrib_area used for converting Q [m3/s] -> q [m/day]

gage_dv           the "main" daily mean flow values used in this .Rmd file 
  _sta            Short name used for this project 
  _Date           Date     
  _waterYear      Year following USGS water year format
  _q1_depth       discharge depth in m/day
  _q7_depth       7-day running average of q_depth
  _q30_depth      30-day running average of q_depth 

pca_input         prcomp() input using gage_dv 


--> 

```{r setup, include=FALSE, message=FALSE}  
#knitr::opts_chunk$set(echo = FALSE)   
options(tibble.print_max = 70) # sets tibble output for printing   

# Sets up the library of packages    
library("here")                   # identifies where to save work  
library("EGRET")                  # Exploration and Graphics for RivEr Trends 
library("rio")                    # more robust I/O - to import and clean data  
library("lubridate")              # easier dates 
library("janitor")                # tools for examining and cleaning dirty data  
library("dataRetrieval")          # USGS data import  
library("measurements")           # eases measurement system manipulation 
library("flextable")              # construct complex table with 'kable'  
library("officer")                # facilitates '.docx' access for table export  
library("forecast")               # for BoxCox.lambda 
library("tidyverse")              # data munging tools 
library("broom")                  # convert statistical objs into tidy tibbles
library("ggfortify")              # data vis tools for statistical analysis 
library("stringi")                # character string processing facilities

# library("mathpix")                # support for 'Mathpix' image to 'LaTeX'  
```

```{r import_daily_flow_metadata, eval=FALSE} 
gage_poss <- whatNWISsites(bBox = c(-103.8, 42.2, -99.2, 44.6),  
                       parameterCd = "00060", 
                       hasDataTypeCd = "dv") %>%  
  arrange(site_no) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# 2. get metadata 
# this needs to be run in parts because some project numbers & 
# inventories are stored as integers & others are stored as characters 

gage_meta_int1 <- gage_poss %>% 
    slice(7, 17, 142, 148) %>% 
  split(.$site_no) %>% 
  map_dfr(~ readNWISsite(siteNumbers = .$site_no)) %>% 
  select(-c(project_no, inventory_dt)) 

gage_meta_int2 <- gage_poss %>% 
    slice(94) %>% 
  split(.$site_no) %>% 
  map_dfr(~ readNWISsite(siteNumbers = .$site_no)) %>% 
  select(-c(project_no, inventory_dt)) 

gage_meta_int3 <- gage_poss %>% 
    slice(-c(7, 17, 94, 142, 148)) %>% 
  split(.$site_no) %>% 
  map_dfr(~ readNWISsite(siteNumbers = .$site_no)) %>% 
  select(-c(project_no, inventory_dt)) 

# 3. join gage metadata 
gage_meta_poss <- bind_rows(gage_meta_int1, 
                            gage_meta_int2, 
                            gage_meta_int3)  

gage_meta_poss <- gage_meta_poss %>% 
  mutate(site_no = zeroPad(site_no, 8)) 

# clean up
rm(gage_meta_int1, gage_meta_int2, gage_meta_int3, gage_poss) 

# 4. export metadata to data folder 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
export(gage_meta_poss, "data/gage_meta_poss.csv")  
```

```{r clean_gage_metadata}
gage_meta_poss <- import("data/gage_meta_poss.csv")  

# 1. remove gages that do not meet standards  
gage_meta <- gage_meta_poss %>%   
  mutate(site_no = as.character(site_no)) %>% 
  mutate(site_no = zeroPad(site_no, 8)) %>%              # pad site_no  
  mutate(reliability_cd = replace_na(reliability_cd, 0))  %>% #  NA <- zero 
  filter(site_no != "06461150"& 
         site_no != "06463670"&  
         site_no != "06461595") %>%  # remove short sites with provisional data 
  filter(site_no != "06441000") %>%  # only active 180 days/yr 
  filter(site_no != "06438000") %>% 
  filter(site_no != "06437000") %>%  # remove northern Black Hills stations 
  filter(site_no != "06442718") %>%  # remove East River stations 
  filter(reliability_cd != "M") %>% # M is minimal data 
  filter(!str_detect(station_nm, 'DAM|DITCH|DRAIN')) %>% # upstream control 
  filter(!str_detect(station_nm, 
                     'CUSTER|KEYSTONE|HILL CITY|HAYWARD')) %>% 
  filter(site_no != "06424000") %>% 
  # crystaline catchments 
  filter(!str_detect(station_nm, 
                     'LEAD|DEADWOOD|WHITEWOOD')) %>%   
  # crystaline catchments 
  filter(!str_detect(station_nm, 'CLEGHORN')) %>% # karstic? spring 
  filter(!str_detect(station_nm, 'BOXELDER|LIME')) %>% # karstic 
  filter(!str_detect(station_nm, 'RAPID')) %>%   
  # Rapid Creek & upper Spring Creek 
  filter(!str_detect(station_nm, 'MISSOURI'))

# 2. remove provisional sites 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
gage_meta <- gage_meta %>% 
  mutate(site_no = as.character(site_no)) %>% 
  select(-reliability_cd) %>% 
  mutate(length = str_length(site_no)) %>% 
  filter(length <= 8) %>%   # removes two provisional sites 
  select(-length)

# 3. remove codes not needed for this project 
gage_meta <- gage_meta %>% 
  select(-agency_cd) %>% 
  select(-site_tp_cd) %>% # all streams so delete 
  select(-c(lat_va, long_va)) %>% # in DMS so delete 
  select(-c(coord_meth_cd, coord_acy_cd)) %>% # coord method & agency, so delete 
  select(-c(coord_datum_cd, dec_coord_datum_cd)) %>% # NAD83 or NAD27 
  select(-c(district_cd, country_cd)) %>% # Congressional dist & Country 
  select(-c(land_net_ds, map_nm, map_scale_fc)) %>%  # refers to USGS maps 
  select(-c(alt_meth_cd, alt_datum_cd, alt_acy_va)) %>% #%>% # altitude metadata  
  select(-c(basin_cd, topo_cd, instruments_cd)) %>% 
  select(-c(construction_dt)) %>% 
  select(-c(tz_cd, local_time_fg)) %>% # daily data, so NA 
  select(-c(gw_file_cd, nat_aqfr_cd, aqfr_type_cd, aqfr_cd)) %>% 
  select(-c(well_depth_va, hole_depth_va, depth_src_cd)) 

rm(gage_meta_poss)
```

```{r add_names_to_gage_meta}

# get data from QGIS analysis of gaged & ungaged watersheds of interest 
wsd_summary <- st_read("sp_data/eco-drought.gpkg", 
                       layer = "wbd_summary", 
                       as_tibble = TRUE) %>% 
  st_drop_geometry() %>% 
  mutate_if(is.factor, as.character) 

# join names to gage_meta  
names <- wsd_summary %>% 
  select(site_no = gage_num, sta = sta_id) %>% 
  drop_na() %>% 
  mutate(site_no = zeroPad(site_no, 8)) 

gage_meta <- full_join(names, gage_meta, 
                     by ="site_no")  

```

```{r import_daily_flow_data} 

# set parameters for Egret::readNWISDaily  
startDate    <- "1979-08-01" # pulling two months early to get Q7 & Q30 
endDate      <- "2018-09-30" 
parameter_cd <- "00060" 

# get daily flows ------------------------------------------------------------ 
# note for EGRET::readNWISDailyQ the discharge is in m^3/s 

# Long River gage needs to be called separately or it creates an error 
gage_lonriv_dv <- gage_meta %>% 
  filter(site_no == "06463500") %>% 
  split(.$site_no) %>% 
  map_dfr(~ readNWISDaily(
    siteNumber = .$site_no, 
    parameter_cd, 
    startDate, 
    endDate), 
    .id = "site_no") 

gage_most_dv <- gage_meta %>% 
  filter(site_no != "06463500") %>%           # drop long river from call 
  split(.$site_no) %>% 
  map_dfr(~ readNWISDaily(
    siteNumber = .$site_no, 
    parameter_cd, 
    startDate, 
    endDate), 
    .id = "site_no") 

# join into a single dataframe, remove add names, & clean up   
gage_dv <-bind_rows(gage_most_dv, gage_lonriv_dv) 

gage_dv <- gage_dv %>% 
  filter(waterYear > "1979")

gage_dv <- full_join(names, gage_dv, 
                     by ="site_no")  

rm(startDate, endDate, parameter_cd)

``` 

```{r filter_short_flow_recs}

# 1. prepare to filter gages with short flow records--------------------------- 
# calculate min & max years of record 
yrs_summary <- gage_dv %>% 
  group_by(site_no) %>%                  
  summarise(years_rec = n_distinct(waterYear),  
            min_year = min(waterYear),  
            max_year = max(waterYear)  
            ) %>%                    
  ungroup() %>% 
  mutate(apparent_yrs = 1 + max_year - min_year) 

gage_meta <- full_join(yrs_summary, gage_meta,      # N = 88 
                       by = "site_no") 
rm(yrs_summary)

# 2. filter gages with less than 6 years of record----------------------------  
# filter gages with no records 
gage_check <- gage_meta %>% 
  filter(is.na(years_rec)) 

gage_meta <- gage_meta %>% 
  filter(!is.na(years_rec))                           # N = 69

# 06400500 CHEYENNE R NEAR HOT SPRINGS SD
# 06403500 FRENCH CR NEAR FAIRBURN SD
# 06404500 BATTLE CR NEAR HERMOSA SD
# 06424500 ELK CR ABOVE PIEDMONT SD
# 06437200 BEAR BUTTE CR NEAR GALENA,SD
# 06437500 BEAR BUTTE CR NEAR STURGIS SD
# 06440500 NORTH FORK BAD R NEAR PHILIP SD
# 06445000 WHITE R BELW COTTONWOOD C N WHITNEY, NEBR.
# 06445500 WHITE R NEAR CHADRON NEBR
# 06446200 WHITE R NEAR ROCKYFORD SD
# 06449250 SPRING CR NEAR ST FRANCIS SD
# 06455900 NIOBRARA RIVER NEAR DUNLAP, NEBR.
# 06456500 NIOBRARA RIVER NR HAY SPRINGS, NEBR.
# 06457000 NIOBRARA RIVER NEAR COLCLESSER, NEBR.
# 06458500 BEAR C NR ELI NEBR
# 06459000 NIOBRARA R NEAR CODY, NEBR.
# 06460900 MINNECHADUZA CREEK NEAR KILGORE, NEBRASKA
# 06463000 NIOBRARA RIVER AT MEADVILLE NE
# 06464000 KEYA PAHA R NEAR HIDDEN TIMBER SD 

# drop stations with less than 6 years of record  
gage_check <- gage_meta %>% 
  filter(years_rec < 6) 

gage_meta <- gage_meta %>% 
  filter(years_rec >= 6)                         # N = 58 

# The following were removed because less than 6 years of record  
# 06400870 HORSEHEAD CR NEAR OELRICHS SD
# 06405400 GRACE COOLIDGE CR NEAR FAIRBURN SD
# 06405500 GRACE COOLIDGE CR NEAR HERMOSA SD
# 06437400 BEAR BUTTE CREEK AT STURGIS, SD
# 06441400 WILLOW CREEK NEAR FORT PIERRE, SD
# 06442130 CEDAR CREEK NR PRESHO, SD
# 06442600 MEDICINE CREEK NR LOWER BRULE, SD
# 06442950 CROW CR NEAR GANN VALLEY SD
# 06445590 BIG BORDEAUX CREEK NEAR CHADRON NEBR
# 06459200 SNAKE RIVER ABV MERRITT RESERVOIR NEBR
# 10150005 Big Beaver Cr at Nebr Hwy 12 nr Valentine, Nebr

# 3. check for incomplete days of record --------------------------------------  
yr_incomp <- gage_dv %>%                          
  group_by(site_no, waterYear) %>%                  
  summarise(days_yr = n()) %>%                    
  ungroup() %>% 
  filter(days_yr < 360) %>% 
  group_by(site_no) %>%   
  summarise(yrs_incomp = n()) %>%    
  ungroup() 

gage_meta <- full_join(yr_incomp, gage_meta,  
                       by = "site_no") %>% 
  filter(!is.na(years_rec))

# 4. remove data with short record between 1990-2015 ------------------------- 

gage_check <- gage_meta %>% 
  filter(is.na(sta))
  
gage_meta <- gage_meta %>% 
  filter(!is.na(sta))

## These didn't make the cut because the original data selection was 
# for 1990-2015  & it would be extremely difficult to put in post-hoc.
# it would be good for the future! 
# Also - its possible to use these for model testing...

# station   min_yr  maxyr  name 
# 06442000   1980   1990   MEDICINE KNOLL CR NEAR BLUNT SD 
# 06459175   1982   1995   SNAKE R AT DOUGHBOY, NE
# 06459500   1980   1995   SNAKE RIVER NEAR BURGE, NEBR. 
# 06454100   1980   1992   Niobrara River at Agate, Nebr. 
# 06461000   1980   1995   MINNECHADUZA CREEK AT VALENTINE, NEBR.
# 06462500   1980   1995   PLUM CREEK AT MEADVILLE, NE
# 06463720   2012   2018   Niobrara River at Mariaville, Nebr. 
# 06400497   1980   1995   CASCADE SPRINGS NEAR HOT SPRINGS SD  
# 06462000   1980   1986   NIOBRARA RIVER NR NORDEN NEBR
# 06463080   1980   1991   LONG PINE CREEK NR LONG PINE, NE
# 06439300   1980   1994   CHEYENNE RIVER AT CHERRY CREEK,SD
# 06442500   1980   1990   MEDICINE CR AT KENNEBEC SD
# 06444000   1980   2004   White River at Crawford, Nebr.
# 06454500   1980   1994   NIOBRARA RIVER ABOVE BOX BUTTE RESERVOIR, NE
# 06455500   1980   1991   NIOBRARA RIVER BELOW BOX BUTTE RESERVOIR NEBR
# 06457500   1980   1991   NIOBRARA RIVER NEAR GORDON, NEBR.

rm(gage_check, yr_incomp, names) 

# sync gage_dv with gage_meta & export results as posthoc to avoid interference 
gage_dv <- semi_join(gage_dv, gage_meta, 
                     by = "site_no")   

export(gage_dv, "data/gage_dv_posthoc.csv") 

# clean up 
rm(names, gage_lonriv_dv, gage_most_dv)
```

```{r prepare-table_gage_metadata}
 
# prepare df for gaging station metadata  
gage_table <- gage_meta %>%  
  mutate(contrib_drain_area_va = 
           coalesce(contrib_drain_area_va, drain_area_va) 
  ) %>% 
  mutate(alt_si = 
           conv_unit(alt_va, "ft", "m")
         ) %>% 
  mutate(contrib_drain_area_si = 
           conv_unit(contrib_drain_area_va, "mi2", "km2")
         ) 

gage_table$years_rec <-  as.integer(gage_table$years_rec) 

gage_table <- gage_table %>% 
select(site_no, station_nm, dec_lat_va, dec_long_va, alt_si, years_rec, 
         contrib_drain_area_si) 

# set column keys for flextable -- these variables also used below 
col_key_num <-gage_table %>% 
  select(dec_lat_va:alt_si, contrib_drain_area_si) %>% 
  names() 

col_key_int <-gage_table %>% 
  select(years_rec, site_no) %>% 
  names() 

# convert tibble to a flextable 
gage_table <- flextable(gage_table) %>% 
  colformat_num(col_keys = col_key_num, 
                big.mark=",", 
                digits = 1, na_str = "N/A") %>% 
  set_header_labels(site_no = "Sta. Number", 
                    station_nm = "Name", 
                    dec_lat_va = "Latitude", 
                    dec_long_va = "Longitude", 
                    alt_si = "Elevation, m", 
                    years_rec = "Years of Record", 
                    contrib_drain_area_si = "Drainage Area, sq-km") %>% 
  autofit() %>% 
  theme_booktabs() 

```

```{r prepare-table_ungaged_metadata}

# prepare df for ungaged watershed metadata  
ung_table <- wsd_summary %>% 
  filter(type == "ungaged") %>%  
  select(HUC12, sta_id, sta_name, dec_lat, dec_lon, cat_out, cat_area)  

# set column keys for flextable 
col_key_num <- ung_table %>% 
  select(dec_lat, dec_lon, cat_out, cat_area) %>% 
  names() 

col_key_int <-ung_table %>% 
  select() %>% 
  names() 

# convert tibble to a flextable 
ung_table <- flextable(ung_table) %>% 
  colformat_num(col_keys = col_key_num, 
                big.mark=",", 
                digits = 1, na_str = "N/A") %>% 
  set_header_labels(HUC12 = "Hydrologic Unit Code", 
                    sta_id = "Station Id", 
                    sta_name = "Name", 
                    dec_lat = "Latitude", 
                    dec_lon = "Longitude", 
                    cat_out = "Elevation, m", 
                    cat_area = "Drainage Area, sq-km") %>% 
  autofit() %>% 
  theme_booktabs() 

```

```{r prepare-table_soils}

# prepare data for soils table 
soils_table <- tibble(
  order = c(6, 5, 4, 3, 2, 1), 
  geo_age = c("Quaternary", "Tertiary", "Tertiary", 
              "Tertiary", "Cretaceous", "Cretaceous"), 
  geo_units = c("Eolean deposits", "Arikaree Group", "Arikaree Group", 
            "White River Group", "Pierre Formation", "Inyan Kara Group"), 
  ecoregion_lvl4	= 
    c("Nebraska Sand Hills", 
      "Keya Paha Tablelands", 
      "Pine Ridge Escarpment", 
      "White River Badlands", 
      "Pierre Shale Plains", 
      "Black Hills Foothills"), 
  soil_order = c("Entisols", 
                 "Entisols, Mollisols", 
                 "Entisols, Mollisols", 
                 "Aridisols, Entisols, Inceptisols", 
                 "Mollisols", 
                 "Alfisols, Entisols"),
  vegetation = 
    c("Sand bluestem, Little bluestem, Prairie sandreed, Big bluestem, Switchgrass", 
      "Blue grama, Sideoats grama, Western wheatgrass, Little bluestem, Needleandthread", 
      "Ponderosa pine, Eastern redcedar, Western snowberry, Skunkbush sumac, Chokecherry, Prairie rose, Little bluestem, Western wheatgrass, Green needlegrass, Prairie sandreed", 
    "Sand sagebrush, Silver sagebrush, Western wheatgrass, Blue gramma, Sideoats grama, Buffalograss", 
    "Little bluestem, Buffalograss", 
    "Ponderosa pine, Little bluestem, Western wheatgrass, Sideoats gramma") 
    ) %>% 
  arrange(order)

# export table for import later 
#export(soils_table, "data/soils_table.csv")

#print.noquote(names(soils_table))

# set column keys for flextable 
col_key_char <- soils_table %>% 
  select(geo_age, geo_units, ecoregion_lvl4, soil_order, vegetation) %>% 
  names() 

# convert tibble to a flextable 
soils_table <- flextable(soils_table) %>% 
   colformat_char(col_keys = col_key_char) %>% 
  set_header_labels(geo_age = "Geologic age", 
                    geo_units = "Major geologic unit", 
                    ecoregion_lvl4	= "Level IV ecoregion", 
                    soil_order = "Soil order", 
                    vegetation = "Major vegetation types") %>% 
  autofit() %>% 
  theme_booktabs()  
```

```{r prepare-table_hydro_metrics}

# prepare df for hydrologic metric table 
hydro_table <- wsd_summary %>% 
  select(type, cat_area, lw_ratio, drain_dens, prcp_mean, t07_mean, vpd_ann, 
         vpd_07, cat_out, cat_rel, slop_med, TWI_mean, perc_cov, fc_mean, 
         ksat_mean, kvert_mean) %>% 
  gather(metric, value, -type) %>% 
  group_by(metric, type) %>% 
  summarize(max = max(value), 
            median = median(value), 
            min = min(value)) %>% 
  ungroup() %>% 
  modify_if(is.numeric, ~round(., digits = 1)) %>% 
  gather(key, stat, -c(type, metric)) %>% 
  unite_("combined", c("key","type")) %>% 
  spread(combined, stat) %>% 
  select(metric, max_gaged, median_gaged, min_gaged, 
         max_ungaged, median_ungaged, min_ungaged) 

# add additional columns 
hydro_tab2 <- tibble(metric = hydro_table$metric, 
                     type = c("planimetric", "elevation", "elevation", 
                              "planimetric", "soils", "soils", "soils", 
                              "planimetric", "cover", "climate", 
                              "elevation", "climate", "elevation", "climate",   
                              "climate"), 
                     transformation = c("1/ln(x)", "sqrt(x)", "sqrt(x)", 
                                        "none", "none", "ln(x)", "ln(x)", 
                                        "ln(x)", "ln(x)", "x^2", "ln(x)", 
                                        "x^2", "x^2", "x^2", "x^2"), 
                     description = c("catchment area", 
                                     "catchment outlet elevation", 
                                     "catchment relief", 
                                     "drainage density", 
                                     "field capacity", 
                                     "forested area (proportion of catchment)", 
                                     "horizontal saturated hydraulic conductivity", 
                                     "vertical saturated hydraulic conductivity", 
                                     "catchment length to width ratio", 
                                     "mean annual precipitation", 
                                     "median catchment slope", 
                                     "mean July temperature", 
                                     "mean topographic wetness index value", 
                                     "July maximum vapor pressure deficit", 
                                     "annual maximum vapor pressure deficit"), 
                     units = c("hectares", "meters", "meters", 
                               "km per square km", "meters per meter", 
                               "sq. meters per sq. meter", "micrometers/sec", 
                               "micrometers/sec", "meters per meter", 
                               "millimeters", "percent", "Centigrade", 
                               "dimensionless", "kilopascals", "kilopascals"), 
                     base_data = c("USGS metadata & feature geometry", "NED", 
                     "NED", "NED", "SSURGO", "NLCD", "SSURGO", "SSURGO", 
                     "feature geometry", "PRISM", "NED", "PRISM", "NED", 
                     "PRISM", "PRISM")
                     ) 

hydro_table <- full_join(hydro_table, hydro_tab2, 
                         by = "metric"
                         )  

hydro_table <- hydro_table %>% 
  select(metric, type, transformation, description, 
         units, base_data, everything()) %>% 
  arrange(type) 

# set column keys & header labels for flextable 
col_key_num <- hydro_table %>% 
  select(-metric) %>% 
  names() 

header_labels <- c(
    metric = "Indicator", 
    type = "Indicator type",       
    transformation = "Transformation", 
    description = "Description",
    units = "Units", 
    base_data = "Base data",    
    max_gaged = "max", 
    median_gaged = "median", 
    min_gaged = "min",      
    max_ungaged = "max", 
    median_ungaged = "median", 
    min_ungaged = "min"
    )    

top_row <- c("", "", "", "", "", "", "Gaged", "Gaged", "Gaged", 
             "Ungaged", "ungaged", "Ungaged")  

hydro_table <- hydro_table %>% 
  flextable() %>% 
  theme_booktabs() %>% 
  colformat_num(col_keys = col_key_num, 
                big.mark=",", 
                digits = 1, na_str = "N/A") %>% 
  set_header_labels(values = header_labels) %>% 
  add_header_row(values = top_row, 
  top = TRUE) %>% 
  merge_at(i = 1, j = 7:9, part = "header") %>% 
  merge_at(i = 1, j = 10:12, part = "header") %>% 
  autofit() 

hydro_table

```

```{r export_flextables_as_docx} 
 
# export a docx of flextables 
tables_ch2 <- read_docx() %>% 
  body_add_flextable(value = gage_table)  %>% 
  body_add_break() %>% 
  body_add_flextable(value = ung_table) %>%  
  body_add_break() %>% 
  body_add_flextable(value = soils_table) %>% 
  body_add_break() %>% 
  body_add_flextable(value = hydro_table) 

print(tables_ch2, target = "output/tables_ch2.docx") 

rm(col_key_char, col_key_int, col_key_num, header_labels, names, top_row, 
   ung_summary)  
rm(gage_table, soils_table, ung_table, tables_ch2)
```

```{r clean_daily_flow, eval=FALSE} 
# checks missing records for Q1, Q7, Q30 

## Q1 missing records 
# "site_no"   "sta"   "Date"    "Q"   "Julian"    
# 06448000 lcr_abv 2016-08-09   NA    60851 
# 06448000 lcr_abv 2016-08-10   NA    60852 

## Q7  missing data (n = 114 ~= 19 x 7) because of start of record for 19 sta  
## Q30 missing data (n = 437 ~= 19 sta x 30 day - 114) 

# 1. using gage_raw to avoid rerunning code above  
#gage_raw <- gage_dv 

# 2. check NA vals for Q1 & remove NA discharge values from record  
gage_check <- gage_dv %>% 
  filter(is.na(Q))  

gage_dv <- gage_dv %>% 
  filter(!is.na(Q)) 

# 3. check NA vals for Q7 & remove NA discharge values from record  
gage_check <- gage_dv %>% 
  filter(is.na(Q7))  

gage_dv <- gage_dv %>% 
  filter(!is.na(Q7)) 

# 3. check NA vals for Q30 & remove NA discharge values from record  
gage_check <- gage_dv %>% 
  filter(is.na(Q30))  

gage_dv <- gage_dv %>% 
  filter(!is.na(Q30))  

# 4. remove provincial data - we removed WY 2018 in the prior analysis 
gage_check <- gage_dv %>% 
  filter(Qualifier != "A") %>% 
  filter(Qualifier != "A:e") %>%  
  filter(Qualifier != "A:<")  

gage_check <- gage_dv %>% 
  filter(waterYear == "2018") 

gage_dv <- gage_dv %>% 
  filter(waterYear != "2018") 

``` 

```{r censor_low_flows, message=FALSE} 
 
# fix low flows ---- this step needed for the PCA  
#   EGRET calculates a "better" zero-flow value, but causes issues with results  
#   this code chunk fixes low flow values by substituting 0.01 cfs 
#   for zero-flow values  
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  

# 1. gather the different flow values 
gage_dv_gath <- gage_dv %>%  # using diff var name to check length later 
  select(site_no, sta, Date, waterYear, Q, Q7, Q30) %>% 
  gather(key = q_type, val = q_val,       # prepares to censor to 0.01 cfs 
         -c(site_no, sta, waterYear, Date)) 

# 2. filter & censor zero flows & non-zero flows -> join flows 
gage_dv_low <- gage_dv_gath %>%  # n ~ 123,987/3 = 41,329 obs or ~7%/yr 
  filter(q_val < 0.01) %>%               
  mutate(q_val = 0.01) 

gage_dv_high <- gage_dv_gath %>% 
  filter(q_val >= 0.01) %>% 
  mutate(q_val = round(.$q_val, digits = 2)) 

gage_dv_gath <- bind_rows(gage_dv_high, gage_dv_low) %>% 
  spread(q_type, q_val) 

# 3. check low-flow stations & results 
gage_dv_low <- gage_dv_low %>% 
  distinct(sta) %>% 
  arrange(sta) 

## Intermittant stations 1990 - 2017 (raw data) 
## bad_fpi, bat_bhr, bat_her, bev_buf, bev_pri, blc_wan, blp_bel, 
## brsf_co, elk_elm, frn_fai, hat_edg ,hor_oel, lcr_bel, plu_hay, 
## ros_ros, spr_her, wcc_ogl, whi_int, whi_kad, whi_ogl, whi_slm 
## whi_sta, whi_whi, wkc_wok  

rm(gage_check, gage_dv_gath, gage_dv_high, gage_dv_low)  

``` 

```{r convert_flow_to_depth, message=FALSE}

# 1. calculate contributing drainage area in square kilometers & sec/day 
gage_contrib_area <- gage_meta %>%  
  select(sta, drain_area_va, contrib_drain_area_va) %>% 
  mutate(contrib_drain_area_va = 
           contrib_drain_area_va %>% 
             is.na %>% 
             ifelse(drain_area_va, contrib_drain_area_va)) %>% 
  select(-drain_area_va) %>% 
  mutate(contrib_drain_km2 = 
           measurements::conv_unit(contrib_drain_area_va, 
                                   from = "mi2",  
                                   to = "km2" 
                                   ) 
         ) %>% 
  mutate(contrib_drain_km2 = 
           round( 
             contrib_drain_km2, 
             digits = 1 
             ) 
         ) %>% 
  mutate(sec_per_day = 60 * 60 * 24) 

# 2. calculate Q-depth:: q = Q/DA [m3/km2] 
gage_dv <- left_join(gage_dv, gage_contrib_area, by = "sta") %>% 
  select(site_no, sta, Date, Q, Q7, Q30, waterYear, 
         contrib_drain_km2, sec_per_day)  %>% 
  gather(type, value, 
         -c(site_no, sta, Date, waterYear, contrib_drain_km2, sec_per_day) 
         ) %>% 
  mutate(value = value * sec_per_day / contrib_drain_km2) %>% 
  spread(type, value) %>% 
  rename(q1_depth = Q) %>% 
  rename(q7_depth = Q7) %>% 
  rename(q30_depth = Q30) %>% 
  select(-c(site_no, contrib_drain_km2, sec_per_day)) 

rm(gage_contrib_area) 
```

```{r PCA_graphs}  

# 1. Prepare to approach normality using forcast::BoxCox --------------------- 

# make dataframes of lambda values
lambda_q1  <- enframe(
  BoxCox.lambda(gage_dv$q1_depth)
  )   

lambda_q7  <- enframe(
  BoxCox.lambda(gage_dv$q7_depth)
  )   

lambda_q30 <- enframe(
  BoxCox.lambda(gage_dv$q30_depth)
  )

# join the dataframes & clean up 
lambda_vals <- bind_cols(lambda_q1, lambda_q7, lambda_q30)  %>% 
  select(-c(name1, name2)) %>% 
  rename(lambda_q1  = value) %>% 
  rename(lambda_q7  = value1) %>% 
  rename(lambda_q30 = value2) 

rm(lambda_q1, lambda_q7, lambda_q30)  

# create a PCA input dataframe 
pca_input <- gage_dv %>% 
  mutate(q1_tr = BoxCox(.$q1_depth, lambda_vals$lambda_q1)) %>% 
  mutate(q7_tr = BoxCox(.$q7_depth, lambda_vals$lambda_q7)) %>% 
  mutate(q30_tr = BoxCox(.$q30_depth, lambda_vals$lambda_q30)) %>% 
  dplyr::select(sta, Date, q1_tr, q7_tr, q30_tr)  

# 2. calculate PCA matrix ----------------------------------------------------- 
pca_matrix <- pca_input %>%  
  select(q1_tr, q7_tr, q30_tr) %>%
  prcomp(., center = TRUE, scale. = TRUE)      

# 3. Gather & summarize PCA results ------------------------------------------
pca_eigen <-tidy(pca_matrix, matrix = "pcs")          # results about PC axes

pca_vars <-  tidy(pca_matrix, matrix = "variables") %>%     # PCA loadings 
  filter(PC != 3) %>%              
  rename(var = column) %>% 
  mutate(PC = as.character(PC)) %>% 
  mutate(pc_stem = "PC") %>% 
  unite("pc_axis", 
        c("pc_stem", "PC"),
        sep = "", 
        remove = TRUE
        ) %>% 
  spread(pc_axis, value)
  
gage_dv <- augment(pca_matrix, data = gage_dv) %>% 
  select(-c(.rownames, .fittedPC3)) %>%       # bind sample vals to PCA matrix 
      mutate(q1_q30_diff = q1_depth - q30_depth) 



```

```{r prepare-plot_add-ecoregions}
# read in shapefile data of intersection of watershed area & 
# merged SD, NE, WY ecoregions  

ecoreg <- st_read("sp_data/ecoreg_int.shp", 
                       as_tibble = TRUE) %>% 
  st_drop_geometry() %>%                       #  we only want the data table 
  mutate_if(is.factor, as.character) %>% 
  rename(ecoreg_L4 = US_L4NAME) %>%    
  rename(ecoreg_L3 = US_L3NAME) %>% 
  rename(ecoreg_L2 = NA_L2NAME) %>% 
  rename(ecoreg_L1 = NA_L1NAME) %>% 
  rename(area_km = Shape_Area) %>% 
  select(sta_id, watshed, ecoreg_L4, ecoreg_L3, 
         ecoreg_L2, ecoreg_L1, area_km) %>% 
# prep for gather -> change title case 
  mutate(grouped_id = row_number()) %>% 
  gather(key, val, -c(sta_id, watshed, grouped_id)) %>%  
  mutate(val = stri_trans_totitle(val)) %>% 
  spread(key, val) %>% 
  select(-grouped_id) %>% 
  mutate(area_km = as.numeric(area_km)  # round the areas 
         ) %>% 
  mutate(area_km = round(area_km, digits = 2)
         )%>%  
# get percentages by ecoregion 
  group_by(sta_id) %>% 
  mutate(area_tot = sum(area_km)) %>% 
  mutate(area_perc = round(
    area_km / area_tot, 
    digits = 2)
  ) %>% 
  ungroup() %>% 
  filter(area_perc > 0.10) %>% 
  arrange(
    desc(area_perc)
    ) %>% 
  arrange(sta_id) %>% 
  select(sta_id, area_perc, ecoreg_L4, 
         ecoreg_L3, ecoreg_L2, ecoreg_L1, watshed)  

# generalize ecoregions using max percentage & percentage near the gage 
ecoreg <- ecoreg %>% 
  group_by(sta_id) %>%  
  arrange(sta_id, desc(area_perc)) %>% 
  # make a default pick based on the max % 
  mutate(ecoreg_wtsd = head(ecoreg_L4, 1) ) %>% 
  # refine the pick based on expert knowlege & percentage near the gage 
  mutate(
    ecoreg_wtsd = 
           case_when(
             sta_id == "bat_bhr" ~ "Semiarid Pierre Shale Plains",  
             sta_id == "blp_bel" ~ "White River Badlands",   
             sta_id == "che_buf" ~ "Semiarid Pierre Shale Plains",  
             sta_id == "che_pla" ~ "Semiarid Pierre Shale Plains",   
             sta_id == "che_red" ~ "Semiarid Pierre Shale Plains",  
             sta_id == "che_sce" ~ "Semiarid Pierre Shale Plains",  
             sta_id == "che_was" ~ "Semiarid Pierre Shale Plains",  
             sta_id == "lcr_bel" ~ "Sand Hills",      
             sta_id == "lwr_aro" ~ "Sand Hills", 
             sta_id == "lwr_mar" ~ "Sand Hills", 
             sta_id == "lwr_ros" ~ "Sand Hills", 
             sta_id == "lwr_vet" ~ "Sand Hills", 
             sta_id == "lwr_whi" ~ "Sand Hills", 
             sta_id == "spr_her" ~ "Black Hills Plateau",   
             sta_id == "whi_int" ~ "White River Badlands",  
             sta_id == "whi_kad" ~ "White River Badlands",  
             sta_id == "whi_oac" ~ "Subhumid Pierre Shale Plains",  
                           TRUE ~ ecoreg_wtsd)
   ) %>% 
  select(sta_id, ecoreg_wtsd, everything()) %>% 
  group_by(sta_id) %>% 
  distinct(ecoreg_wtsd, .keep_all = TRUE) %>% 
  # generalize the Pierre shale 
  mutate(
    ecoreg_wtsd = 
           case_when(
             ecoreg_wtsd == 
               "Semiarid Pierre Shale Plains" ~ "Pierre Shale Plains", 
             ecoreg_wtsd == 
               "Subhumid Pierre Shale Plains" ~ "Pierre Shale Plains", 
                           TRUE ~ ecoreg_wtsd)
   ) %>% 
  arrange(ecoreg_wtsd) 

# join ecoregions to gage_dv 
eco_join <- ecoreg %>% 
  select(sta = sta_id, ecoreg = ecoreg_wtsd) 

gage_dv <- full_join(gage_dv, eco_join, 
                  by = "sta")


#  4. Create summaries of the results ----------------------------------------

pca_mon <- gage_dv %>% 
  mutate(mon = month(Date)) %>% 
  group_by(sta, mon, waterYear) %>% 
  summarize(PC1_mean    = mean(.fittedPC1),
            PC2_mean    = mean(.fittedPC2), 
            q1_mean     = mean(q1_depth), 
            q7_mean     = mean(q7_depth),
            q30_mean    = mean(q30_depth), 
            q1_q30_mean = mean(q1_q30_diff),  
            ecoreg      = first(ecoreg)
            ) %>% 
  ungroup() %>% 
  arrange(PC2_mean) %>% 
  arrange(PC1_mean) 

pca_ann <- gage_dv %>% 
  group_by(sta, waterYear) %>% 
  summarize(PC1_mean    = mean(.fittedPC1),
            PC2_mean    = mean(.fittedPC2), 
            q1_mean     = mean(q1_depth), 
            q7_mean     = mean(q7_depth),
            q30_mean    = mean(q30_depth), 
            q1_q30_mean = mean(q1_q30_diff),  
            ecoreg      = first(ecoreg)
            ) %>% 
  ungroup() %>% 
  arrange(PC2_mean) %>% 
  arrange(PC1_mean) 

pca_sum <- gage_dv %>% 
  group_by(sta) %>% 
  summarize(PC1_mean    = mean(.fittedPC1),
            PC2_mean    = mean(.fittedPC2), 
            q1_mean     = mean(q1_depth), 
            q7_mean     = mean(q7_depth),
            q30_mean    = mean(q30_depth), 
            q1_q30_mean = mean(q1_q30_diff),  
            ecoreg      = first(ecoreg)
            ) %>% 
  ungroup() %>% 
  arrange(PC2_mean) %>% 
  arrange(PC1_mean) 
```


```{r pc-plot} 



p1 <- autoplot(pca_matrix, 
               data = test, 
               scale = 0, 
               colour = "ecoreg", 
               size = 0.1, 
               alpha = 0.5) 
p1 + theme_bw()


ggplot(pca_sum, aes(PC1_mean, PC2_mean)) + 
  geom_jitter(
    aes(
      shape = ecoreg, 
      colour = ecoreg), 
    size = 1.5
    ) + 
  theme_classic()


```


# prepare data 
#pca_input <- gage_input %>%  
#  select(date, sta, q, q7, q30, contrib_drain_area_va)

# training data: 
# num  lambda_q1  lambda_q7  lambda_q30 PC1perc PC2perc PC1cum  PC2cum 
#  01    0.013      0.112      0.265     0.942   0.050   0.942   0.992  
#  02   -0.005      0.070      0.223     0.949   0.047   0.949   0.994    
#  03   -0.102     -0.136     -0.148     0.959   0.035   0.959   0.994   
#  04   -0.008      0.071      0.242     0.947   0.047   0.947   0.994   
#  05   -0.056     -0.113     -0.187     0.953   0.041   0.953   0.994 
#  06   -0.031     -0.070     -0.120     0.960   0.034   0.960   0.994 
#  07   -0.055     -0.067     -0.060     0.966   0.029   0.966   0.995 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# vars:
# run  PC  q1_depth  q7_depth  q30_depth       
#   01  1    0.577      0.589     0.566   
#   01  2    0.584      0.186    -0.790     
#   02  1    0.578      0.588     0.567       
#   02  2    0.567      0.210    -0.796      
#   03  1    0.578      0.584     0.569           
#   03  2    0.545      0.243    -0.803       
#   04  1    0.578      0.588     0.566           
#   04  2    0.571      0.205    -0.795       
#   05  1    0.578      0.586     0.568           
#   05  2    0.555      0.228    -0.800       
#   06  1    0.578      0.585     0.569            
#   06  2    0.553      0.233    -0.800       
#   06  1    0.578      0.583     0.571             
#   06  2    0.542      0.248    -0.803  
#   07
#   07
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 

#Mclust model:
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
# num  model  log.likeli  n  df   BIC     ICL     purpose 
#  01  VEV     116.42    26  25  151.4   151.0   estimate NA 
#  02  VEV      82.65    23  17  112.0   111.6   estimate NA 
#  03  VVV     139.30    29  19  214.6   214.5   estimate NA   
#  04  EEV     161.53    24  65  116.5   116.4   estimate NA    
#  05  EEE     139.33    24  37  161.6   161.0   estimate NA    
#  06  EEV     130.93    27  19  199.2   199.2   estimate NA    
#  07  VVV     151.58    31  19  237.9   237.7   estimate NA    
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
# run#  Cl1#  Cl2#  Cl3#  Cl4#  Cl5#  Cl6#  Cl7#    Cl8#    Cl9#
#  01   14      8     4    na    na    na    na      na      na  
#  02   15      8    na    na    na    na    na      na      na  
#  03   14     15    na    na    na    na    na      na      na  
#  04    3      4     3     2     4     2     1       1       4    
#  05    2      9     3     1     5     2     1       1      na  
#  06   14     13    na    na    na    na    na      na      na  
#  07    7     24    na    na    na    na    na      na      na  
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 

# a final cluster analysis was conducted after missing values were found 
# two runs for final cluster analysis are conducted 

# | Run_num | description | num_obs | 
# |:-------:|:-----------:|:-------:|
# |    1    | all years   | 317,036 |                                     
# |    2    | wet years   | 147,188 |
# |    2    | dry years   | 169,848 |

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# the series of station runs are as follows: 

pca_eigen 
# num  lambda_q1  lambda_q7  lambda_q30 PC1perc PC2perc PC1cum  PC2cum 
# all   -0.082     -0.086     -0.027     0.967   0.028   0.967   0.995 
# wet   -0.060     -0.078     -0.089     0.963   0.031   0.963   0.994 
# dry   -0.145     -0.185     -0.211     0.963   0.032   0.963   0.995
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 

pca_vars 
# run  PC  q1_depth  q7_depth  q30_depth   notes    
#  all  1    0.578      0.583     0.571           
#  wet  1    0.578      0.584     0.570           
#  dry  1   -0.578     -0.584    -0.570   need to inverse
#  all  2    0.540      0.251    -0.803       
#  wet  2    0.546      0.243    -0.802    
#  dry  2   -0.546     -0.243     0.802   need to inverse   

# eigenvecter:
#   96% of covarience is explained by PCA1 & 4% of varience by PCA2
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
# variables:
#   PC1 - approximately equal loadings of Q1, Q7, Q30; 
#         approximates hydrologic export 
#   PC2 - large opposite loadings of Q1 & Q30  
#         contribution of baseflow vs event-flow
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Preparing to cluster by the central tendancy
gage_pca_sum <- gage_pca_sum %>% 
  arrange(sta)

gage_clust <- gage_pca_sum %>% 
  select(q1_mean:q30_mean)

gage_clust_meta <- gage_pca_sum %>% 
  select(sta)

# apply model-based-clustering, extract results & add metadata
gage_clust_l <- Mclust(gage_clust) 

gage_clust <- as_tibble(gage_clust_l$z) 
gage_clust <- bind_cols(gage_clust_meta, gage_clust) 

# drop low probs - join & rearrange 
gage_clust <- gage_clust %>% 
  gather(key = group, value = prob, -sta) %>% 
  filter(prob > 0.5) 

gage_clust <- full_join(gage_pca_sum, gage_clust, 
                             by = "sta")

gage_clust <- gage_clust %>% 
  dplyr::select(group, sta, everything()) %>% 
  arrange(sta) %>%  
  arrange(group)

# join cluster to PCA data & arrange
gage_mod <- full_join(gage_pca, gage_clust, by = "sta") 

gage_mod <- gage_mod %>% 
  select(group, sta, everything()) %>% 
  select(-c(q:q30_tr)) %>% 
  select(-eigen_dist) 

summary(gage_clust_l) # Print a summary 

#clean up environment 
rm(pca_input, pca_matrix, pca_check, pca_eigen, pca_vars) 
rm(gage_clust_meta, gage_pca_sum, gage_pca, gage_input) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#Mclust model:

# num  model  log.likeli  n  df   BIC     ICL     
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# all  VEV     190.89    31  41  241.0   240.8  
# wet  EEE     175.15    31  29  250.7   249.4  
# dry  EEE     165.01    31  37  203.0   202.6   
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Clustering table:
# run#  Cl1#  Cl2#  Cl3#  Cl4#  Cl5#  Cl6#  Cl7#    Cl8#  
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# all    7     11     5     3     5    na    na      na     
# wet    3      5     1     4    16     2    na      na    
# dry    4     12     5     1     1     6     1       1     
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Model-based clustering selected a model with two-three components 
# (i.e. clusters). The optimal selected model name is VEV model. 
# That is the five components are ellipsoidal with varying volume, 
# and orientation, and equal shape. The summary contains also the 
# clustering table specifying the number of observations in each 
# clusters.

tidy(pca_matrix, matrix = "u")

test <- augment(pca_matrix, pca_input)
```

