```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE, message=FALSE}
# Sets up the library of packages 
library("tidyverse")
library("here") # identifies where to save work
library("rio") # more robust I/O - to import and clean data
library("janitor") # tools for examining and cleaning dirty data
library("DataExplorer")
library("rnoaa") #  R wrapper for NOAA data inc. NCDC
library("lubridate")

# library("dataRetrieval") # USGS data import
# library("RColorBrewer") - there is a better one?
# library("workflowr") # creates a research website
# library("colorspace")
# library("bookdown") # 
# library(unpivotr) # fix nasty Excel files
# library("friendlyeval")

# a useful description of commits:
# http://r-pkgs.had.co.nz/git.html
```


## Overview
<!--This is the chapter 2 hydrologic drought.  
Next Steps:
1. Methods.
2. Find the ramblings I wrote about scope and purpose
3. figure out how to reference stuff 
-->

## Introduction


## Methods
I imported Global Historical Climatology Network (GHCN) daily precipitation records for candidate "WEATHER STATIONS" into R-Studio (REF1) using the "rnoaa" package.
I used Theissen polygons and the length and continuity of precipitation records to select stations for further analysis.
I used 'dplyr' to fill NA values with data from nearest station
I used 'dplyr' to create monthly vals from daily vals.
I removed short records: Oral & Long Valley after checking for 
covariance.


